### 2023-11-24 co687/project/main.tex

\documentclass[letterpaper,12pt]{article}

\usepackage{amsmath,amssymb,amstext}
\setlength{\marginparwidth}{0pt} % width of margin notes
% N.B. If margin notes are used, you must adjust \textwidth, \marginparwidth
% and \marginparsep so that the space left between the margin notes and page
% edge is less than 15 mm (0.6 in.)
\setlength{\marginparsep}{0pt} % width of space between body text and margin notes
\setlength{\evensidemargin}{0.125in} % Adds 1/8 in. to binding side of all 
% even-numbered pages when the "twoside" printing option is selected
\setlength{\oddsidemargin}{0.125in} % Adds 1/8 in. to the left of all pages when "oneside" printing is selected, and to the left of all odd-numbered pages when "twoside" printing is selected
\setlength{\textwidth}{6.375in} % assuming US letter paper (8.5 in. x 11 in.) and side margins as above

\begin{document}

\input{00-introduction.tex}
\input{01-preliminaries.tex}
\input{02-basis-reduction.tex}
\input{03-nearest-plane-cvp.tex}
\input{04-ggh-trapdoor.tex}

\end{document}