### 2023-11-24 co687/project/main.tex
Some new formatting

\documentclass[letterpaper,12pt]{article}
\usepackage{amsmath,amssymb,amstext}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amssymb}

\title{A introduction to lattice trapdoor}
\author{Ganyu Xu (g66xu)}
\date{Fall, 2023}

\begin{document}
\maketitle

\input{00-introduction.tex}
\input{01-preliminaries.tex}
\input{02-basis-reduction.tex}
\input{03-nearest-plane-cvp.tex}
\input{04-ggh-trapdoor.tex}

\end{document}