### 2024-02-02 co789/homework-02/q4/main.tex

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 4}

\subsection*{(a)}
By the definition of the (quotient) ring $R_q = \mathbb{Z}_q[x] / \langle p(x) \rangle$ we know that:

\begin{equation}
    a(x)s(x) + e(x) = p(x)g(x) + b(x)
\end{equation}

Where $g(x)$ is some polynomial in $\mathbb{Z}_q[x]$.

Where $\omega$ is a root of $p(x)$, evaluating equation (1) at $x = \omega$ is as follows:

$$
a(\omega)s(\omega) + e(\omega) = 0 \cdot g(\omega) + b(\omega) = b(\omega)
$$

$\blacksquare$

\subsection*{(b)}
From equation (1) we know that $a(\omega)s(\omega) + e(\omega) = b(\omega)$ if and only if $p(\omega)g(\omega) = 0$. Where $\omega$ is not a root of $p(x)$, the equality holds if and only if $g(\omega) = 0$.

While $g(\omega) = 0$ does not hold in general, in specific cases it can still happen. For example, if the sum of degrees of $a(x)$ and $s(x)$ is less than the degree of $p(x)$, then $g(x) = 0$ (aka $b(x)$ does not need to be reduced modulus $p(x)$).

\end{document}