### 2024-02-21 co789/homework-03/q5/main.tex

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Question 5}

\subsection*{1}
For a given message $m$, we can forge a signature under the public key $(A, \mathbf{t})$ using the following procedure:

\begin{enumerate}
    \item Sample $\mathbf{y} \leftarrow \chi_y$ 
    \item Compute $\mathbf{w} \leftarrow A\mathbf{y}$
    \item Compute $c \leftarrow H(\mathbf{w}, \mathbf{t})$
    \item Give $(A, \mathbf{w} + c\mathbf{t})$ to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ solver, which returns some $\mathbf{z}$
    \item Output $\sigma = (\mathbf{w}, c, \mathbf{z})$
\end{enumerate}

$\sigma$ is a valid forgery because $\mathbf{z}$ as a solution to the $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1 - np\tau)$ problem satisfies the verification conditions:

\begin{enumerate}
    \item $\norm{\mathbf{z}}_\infty \leq \gamma_1 - np\tau$
    \item $A\mathbf{z} = \mathbf{w} + c\mathbf{t}$
\end{enumerate}

\subsection*{2}
The key recovery attack is as follows:

\begin{enumerate}
    \item Sample some random message $m$
    \item Query the signature of $m$, which is $(\mathbf{w}, c, \mathbf{z})$
    \item Give $(A, \mathbf{w})$ to $\operatorname{module-ISIS}(k, l, q, p(x), \gamma_1)$ solver, which returns some $\mathbf{y}$
    \item Compute $\mathbf{s} = c^{-1}(\mathbf{z} - \mathbf{y})$. $\mathbf{s}$ is the secret key
\end{enumerate}

This procesure works because for the queried signature to be valid, it must satisfy $\mathbf{z} = \mathbf{y} + c\mathbf{s}$, meaning that if we can recover $\mathbf{y}$, then we can recover $\mathbf{s}$. In this instance, a valid $\mathbf{y}$ is recovered using the solver on $A\mathbf{y} = \mathbf{w}$.

\end{document}