### 2024-03-10 co789/homework-04/q6/main.tex

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage{algorithm}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 6}
We will prove equivalence by showing that each statement implies the other.

If (1) is true for all $\mathbf{c}_1 \neq \mathbf{c}_2$, then we can fix $\mathbf{c}_2 = \mathbf{0}$, and statement (1) becomes

$$
\norm{\mathbf{c}_1 - \mathbf{0}}_H = \norm{\mathbf{c}_1}_H \geq d
$$

for all non-zero $\mathbf{c}_1$, thus statement (2) is true.

Now suppose (2) is true. For all $\mathbf{c}_2 \neq \mathbf{c}_1$, $\mathbf{c}_2 - \mathbf{c}_1$ is a non-zero code in $C$ because $C$ is linear. Therefore:

$$
\norm{\mathbf{c}_2 - \mathbf{c}_1}_H \geq d
$$

thus statement (1) is true.

\end{document}