### 2024-03-30 co789/homework-05/q1/main.tex
Drafted solution to problem 1, 2, 4

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\metric}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 1}

We will prove $\metric{c_1 - (c_2 + e)} \geq \metric{c_1 - c_2} - \metric{e}$ by proving a more general result:

\begin{theorem}[variation of triangle inequality]
    let $\vert \cdot \vert$ be a metric, then for $a, b$ in the vector space:

    \begin{equation*}
        \metric{a - b} \geq \metric{a} - \metric{b}
    \end{equation*}
\end{theorem}

The desired result is a special case of the theorem above where $a = c_1 - c_2$ and $b = e$

\begin{proof}
    By the triangle inequality we know that $\metric{a + b} \leq \metric{a} + \metric{b}$. Let $c = a + b$, then $a = c - b$, and the triangle inequality becomes

    \begin{equation*}
        \metric{c} \leq \metric{c - b} + \metric{b}
    \end{equation*}

    Re-arranging the equation above gives us the following:

    \begin{equation*}
        \metric{c - b} \geq \metric{c} - \metric{b}
    \end{equation*}

    We made no assumption of $a$, $b$, or $c$, so the inequality above holds for all vectors $b, c$ in the vector space
\end{proof}

\end{document}