### 2024-04-02 information-theory/main.tex
Getting started with ECE612 cheatsheet

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\Norm}[1]{\Vert {#1} \Vert}
\newcommand{\norm}[1]{\vert {#1} \vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{ECE 612, Information Theory}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Winter, 2024}

\begin{document}
%%%% TITLE %%%%%
\maketitle

\section*{Preliminares}
\begin{definition}
    The normal distribution $N(\mu, \sigma^2)$ has the probability density function

    $$
    f(x) = \frac{1}{\sigma\sqrt{2\pi}}\exp(-\frac{1}{2}(\frac{x-\mu}{\sigma})^2)
    $$
\end{definition}

\begin{definition}
    The joint normal distribution $N(\mathbf{\mu}, K)$ is defined by probability density function:

    $$
    f(\mathbf{x}) = \frac{1}{\sqrt{(2\pi)^n\det(K)}}
    \exp(-\frac{1}{2}
        (\mathbf{x} - \mathbf{\mu})^\intercal
        K^{-1}
        (\mathbf{x} - \mathbf{\mu})
    )
    $$
\end{definition}

\section{Entropy, mutual information, divergence}

\section{Entropy rate}

\section{Asymptotic equipartition property}

\section{Data compressions}

\section{Channel capacity}

\section{Differential entropy}
\begin{theorem}[Differential entropy of Gaussian distribution]
    Let $X$ be Gaussian $N(0, \sigma^2)$, then

    $$
    h(X) = \frac{1}{2}\log{(2\pi e \sigma^2)}
    $$
\end{theorem}

\begin{theorem}
    Let $\mathbf{X}$ follow joint Gaussian distribution $N(\mathbf{0}, K)$, then:

    \begin{equation*}
        h(\mathbf{X}) = \frac{1}{2}\log ((2\pi e)^n \det{K})
    \end{equation*}
\end{theorem}

\section{Gaussian channel}
\begin{definition}[Gaussian channel with power constraint]

\end{definition}

\begin{definition}[Information channel capacity]

\end{definition}

\begin{theorem}
    The information channel capacity of a Gaussian channel is

    \begin{equation}
        \max_{f_X: E[X^2] \leq P} I(X; Y) = \frac{1}{2}\log{(1 + \frac{P}{\sigma^2})}
    \end{equation}

    Where $P$ is the power constraint, and $\sigma^2$ is the variance of the Gaussian noise. The maximum is achieved when $X$ follows Gaussian distribution $X \leftsample N(0, P)$
\end{theorem}

\section{Rate distortion theory}
\end{document}