### 2024-07-14 archives/co789/homework-01/q02/main.tex

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For code
\usepackage{listings}

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
% Title is not required when submitting to Crowdmark
% \maketitle

\section*{Q2}
Where $m \leq n$, if $A \in \mathbb{F}_q^{m \times n}$ is full rank, then the columns of $A$ span $\mathbb{F}_q^m$. Since $\mathbf{s} \leftarrow \mathbb{Z}_q^n$ is uniformly sampled, it naturally follows that $A\mathbf{s}$ is uniformly distributed across $\mathbb{F}_q^m$. As a result, $\mathbf{b} \leftarrow A\mathbf{s} + \mathbf{e}$ is uniformly distributed across $\mathbb{F}_q^m$ regardless of the choice of distribution of $\mathbf{e} \leftarrow \chi_e^m$, and the LWE samples are identically distributed as uniform random noise. Thus, when $m \leq n$, $A$ is full rank, $\mathbf{s}$ is uniformly random, decisional-LWE is information-theoretically hard.

P.S. Given two independent random variables $A, B \in G$ whose support lies in some finite group, if $A$ is uniformly random, then $A + B$ is also uniformly random regardless of the distribution of $B$:

$$
\begin{aligned}
P(A + B = c)
&= \sum_{a \in G}P(A=a)P(B=c - a) \\
&= \sum_{a \in G}\frac{1}{\vert G \vert}P(B=c - a) \\
&= \frac{1}{\vert G \vert}\sum_{a \in G}P(B=c - a) \\
&= \frac{1}{\vert G \vert} \cdot 1 = \frac{1}{\vert G \vert} \\
\end{aligned}
$$

\end{document}