### 2024-07-14 archives/co789/homework-05/q7/main.tex

\documentclass{article}
\usepackage[margin=1in,letterpaper]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}

% For source code
\usepackage{listings}

% Algorithms and pseudocode
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{algpseudocode}

% Common shortcuts
\newcommand{\round}[1]{\lfloor {#1} \rceil}
\newcommand{\norm}[1]{\Vert {#1} \Vert}
\newcommand{\var}[1]{\operatorname{Var}[{#1}]}
\newcommand{\leftsample}{\overset{{\scriptscriptstyle\$}}{\leftarrow}}

% Environments: definitions, theorems, propositions, corollaries, lemmas
%    Theorems, propositions, and definitions are numbered within the section
%    Corollaries are numbered within the theorem, though they are rarely used
\newtheorem{definition}{Definition}[section]
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{remark}{Remark}
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{proposition}{Proposition}[section]
\newtheorem{lemma}{Lemma}[theorem]

\title{CO 789, Homework 1}
\author{Ganyu (Bruce) Xu (g66xu)}
\date{Fall 2023}

\begin{document}
%%%% TITLE %%%%%
% \maketitle

\section*{Question 7}
Recall the Beaver triple protocol for computing $xy$ given $[x], [y]$:

\begin{enumerate}
    \item A trusted third party generates random $a, b \leftsample \mathbb{Z}_n^\ast$ and compute $c = ab$
    \item This third party distributes $[a], [b], [c]$ to the computation parties. We claim that since the circuit of the computation is known ahead of time, the trusted parties can generate and distributes $[a], [b], [c]$ before the MPC actually starts. Therefore, distributing $[a], [b], [c]$ does not count toward online communication cost during the MPC
    \item Each party computes $[x-a], [y-b]$, broadcasts their shares, which publicizes $\epsilon = x-a, \delta = y-b$. \textbf{This invokes 2 units of online communication}
    \item Each party computes $[z] = \delta[x] + \epsilon[y] - \epsilon\delta + [c]$ which evaluates to $[xy]$
    \item Each party opens $[z]$, which assembles back into $xy$
\end{enumerate}

When computing the square of a number, we don't need to obfuscate two distinct operands $x, y$; instead, we only need to obfuscate one operand $x$, so we only need to generate a single one-time pad $a$ and its square $c = a^2$. This means that individual parties only need to broadcast $[x-a]$ during their computation, bring the units of online communication during MPC from 2 to 1.

Here is the full protocol

\begin{enumerate}
    \item A trusted third party generates random $a \leftsample \mathbb{Z}_n^\ast$ and computes $c = a^2$
    \item The shares $[a], [c]$ are pre-distributed to the computation parties.
    \item Each party computes $[x-a]$ and boardcasts the value of their share, which publicizes the value of $x-a$. \textbf{This invokes 1 unit of online communication}.
    \item Each party computes $[z] = 2\cdot (x-a)[x] - (x-a)^2 + [c]$, which evaluates to $[x^2]$
    \item Each party opens $[z]$, which assembles back into $x^2$
\end{enumerate}

Proof of step 4:

$$
\begin{aligned}
    \lbrack z \rbrack &= 2\cdot (x-a)[x] - (x-a)^2 + [c] \\
    &= [2x^2 - 2ax] - (x^2 - 2ax + a^2) + [a^2] \\
    &= [2x^2 - 2ax - (x^2 - 2ax + a^2)] + [a^2] \\
    &= [x^2 - a^2] + [a^2] \\
    &= [x^2]
\end{aligned}
$$

\end{document}