### 2024-10-11 etm-transform/manuscript/main.tex
Adding revisions

%%%% 1. DOCUMENTCLASS %%%%
%
% \documentclass[journal=tches,submission,notanonymous]{iacrtrans}
\documentclass[journal=tches,submission]{iacrtrans}
%%%% NOTES:
% - Change "journal=tosc" to "journal=tches" if needed
% - Change "submission" to "final" for final version
% - Add "notanonymous" to reveal authors
% - Add "spthm" for LNCS-like theorems

%%%% 2. PACKAGES %%%%
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{caption}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{positioning}
% \usepackage[demo]{graphicx}

%%%% 3. AUTHOR, INSTITUTE %%%%
\author{
    Ganyu Xu\inst{1}
    \and Kalikinkar Mandal\inst{2}
    \and Guang Gong\inst{1}
}
\institute{
  University of Waterloo, Waterloo, Canada, \email{{g66xu,ggong}@uwaterloo.ca}
  \and
  University of New Brunswick, New Brunswick, Canada, \email{kmandal@unb.ca}
}
%%%% NOTES:
% - We need a city name for indexation purpose, even if it is redundant
%   (eg: University of Atlantis, Atlantis, Atlantis)
% - \inst{} can be omitted if there is a single institute,
%   or exactly one institute per author

% Custom commands
\newcommand{\pke}{\texttt{PKE}}
\newcommand{\keygen}{\texttt{KeyGen}}
\newcommand{\encrypt}{\texttt{Enc}}
\newcommand{\decrypt}{\texttt{Dec}}
\newcommand{\kem}{\texttt{KEM}}
\newcommand{\encap}{\texttt{Encap}}
\newcommand{\decap}{\texttt{Decap}}
\newcommand{\etm}{\texttt{EtM}}  % encrypt-then-mac
\newcommand{\mac}{\texttt{MAC}}
\newcommand{\sign}{\texttt{Sign}}
\newcommand{\verify}{\texttt{Verify}}
\newcommand{\pk}{\texttt{pk}}
\newcommand{\sk}{\texttt{sk}}
\newcommand{\pco}{\texttt{PCO}}
\newcommand{\cvo}{\texttt{CVO}}
\newcommand{\leftsample}{\stackrel{\$}{\leftarrow}}
\newcommand{\llbrack}{[\![}
\newcommand{\rrbrack}{]\!]}
\newcommand{\norm}[1]{\left\lvert #1 \right\rvert}
\newcommand{\adv}{\texttt{Adv}}
\newcommand{\fotplus}{\texttt{FOT+}}
\newcommand{\us}{\mu s}
\def\mlkemplus{\text{ML-KEM}^+}

%%%% 4. TITLE %%%%
\title{Faster generic IND-CCA secure KEM using encrypt-then-MAC}
%%%% NOTES:
% - If the title is too long, or includes special macro, please
%   provide a "running title" as optional argument: \title[Short]{Long}
% - You can provide an optional subtitle with \subtitle.

\begin{document}

\maketitle

%%%% 5. KEYWORDS %%%%
\keywords{
    Key encapsulation mechanism, 
    message authentication code,
    post-quantum cryptography,
    lattice cryptography,
    Fujisaki-Okamoto transformation
}

%%%% 6. ABSTRACT %%%%
\begin{abstract}
    Indistinguishability under chosen-ciphertext attack (IND-CCA) is the accepted standard of security for key encapsulation mechanisms (KEM). However, provably chosen-ciphertext security is difficult to achieve from scratch. In this work, we generalize the ``Hashed ElGamal'' encryption scheme proposed by M. Abdalla et al into an IND-CCA secure KEM by combining an OW-PCA secure public-key encryption (PKE) scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC. Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC transformation replaces de-randomization and re-encryption with computing a MAC tag. When instantiated with the PKE sub-routines of ML-KEM, the encrypt-then-MAC transformation achieves massive computational efficiency improvements over ML-KEM in a variety of practical scenarios.
\end{abstract}

\section{Introduction}\label{sec:introduction}
A key encapsulation mechanism (KEM) \cite{DBLP:journals/iacr/Shoup01} is a cryptographic primitive that allows two parties to establish a shared secret over an insecure channel. The desired security standard for a KEM is called indistinguishability under chosen-ciphertext attack (IND-CCA). Intuitively speaking, IND-CCA security requires that no efficient adversary can distinguish a pseudorandom shared secret from a uniformly random bit string of equal length, even with access to a decapsulation oracle throughout the attack. However, building a provably IND-CCA secure KEM is tremendously difficult. Early attempts without formal proofs, such as RSA encryption defined in PKCS\#1 v1.5 \cite{DBLP:journals/rfc/rfc2313}, were later shown to be vulnerable to practical chosen-ciphertext attacks \cite{DBLP:conf/crypto/Bleichenbacher98}. In recent decades, the most viable approach has been to start with cryptographic primitives possessing weaker security properties, such as a public-key encryption (PKE) scheme with one-way security under chosen-plaintext attack (OW-CPA), then add steps to achieve \emph{ciphertext non-malleability} \cite{DBLP:conf/asiacrypt/BellareN00}. Some of the earliest proposals for generic IND-CCA secure constructions include OAEP \cite{DBLP:conf/eurocrypt/BellareR94}, Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13}, REACT \cite{DBLP:conf/ctrsa/OkamotoP01}, and GEM \cite{DBLP:conf/ctrsa/CoronHJPPT02}.

On the other hand, chosen-ciphertext security is a solved problem in symmetric cryptography. It is well understood that, by combining an IND-CPA secure symmetric encryption scheme with an existentially unforgeable message authentication code (MAC) in a pattern called encrypt-then-MAC \cite{DBLP:conf/crypto/Krawczyk01}, one can build an authenticated encryption scheme \cite{DBLP:conf/asiacrypt/BellareN00} that achieves IND-CCA security. While this technique cannot be directly applied in the context of public-key cryptography due to the lack of a shared symmetric key between the two communicating parties, the concept of authenticating ciphertext using a MAC still has strong merits. Abdalla, Rogaway, and Bellare proposed DHIES (also known as ``Hashed ElGamal'')\cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}, a hybrid public-key encryption (HPKE) scheme whose IND-CCA security reduces to the Gap Diffie-Hellman assumption \cite{DBLP:conf/pkc/OkamotoP01} under the random oracle model. The technique behind DHIES is to derive both the shared secret and a symmetric MAC key by hashing a random PKE plaintext, encrypting the PKE plaintext, then authenticating the PKE ciphertext using the previously derived MAC key. When the Gap Diffie-Hellman assumption holds and the MAC is existentially unforgeable, no efficient adversary can recover the decryption of an unknown ciphertext even with access to a decryption oracle because it cannot produce a valid tag for such unknown ciphertext.

\subsection{Our contribution}
In this paper, we propose a generic IND-CCA secure KEM using techniques adapted from DHIES. We call our proposed scheme the encrypt-then-MAC transformation due to the conceptual similarity to the namesake symmetric encryption technique for achieving authenticated encryption. Our scheme mainly differs from DHIES in its versatility and input requirement. Whereas the IND-CCA security of DHIES reduces specifically to the Gap Diffie-Hellman assumption, the chosen-ciphertext security of the encrypt-then-MAC KEM reduces more generally to the OW-PCA security \cite{DBLP:conf/ctrsa/OkamotoP01} of the input scheme (we will show in section \ref{sec:main-results} that the Gap Diffie-Hellman assumption implies a special case of OW-PCA security). In addition, we propose that because each call to encapsulation samples a fresh PKE plaintext, the encrypt-then-MAC KEM can be instantiated with one-time secure MAC such as Poly1305 (Abdalla et al originally proposed to use HMAC and CBC-MAC, which are many-time secure MAC but less efficient than one-time MAC, see section \ref{sec:choosing-a-message-authenticator}), which greatly improves the computational efficiency of the scheme.

\begin{figure}
    \centering
    \begin{tikzpicture}
        \tikzstyle{rect} = [draw, rectangle, minimum width=3cm, minimum height=1cm]
        \tikzstyle{filledrect} = [draw, rectangle, minimum width=3cm, minimum height=1cm, fill=lightgray]
        \matrix [column sep=7mm, row sep=5mm] {
            \node (pk) [rect] {$\pk$}; &
            \node (m) [rect] {$m \leftsample \mathcal{M}$}; &
            \node (hashg) [rect] {Hash function $G$}; \\
            \node (pkeenc) [rect] {$\texttt{PKE.Enc}$}; &
            \node (mackey) [rect] {$k_\mac$}; \\
            \node (ct) [filledrect] {$c^\prime$}; &
            \node (mac) [rect] {$\mac$}; &
            \node (tag) [filledrect] {$t$}; \\
        };
        \draw[->, thick] (pk) -- (pkeenc);
        \draw[->, thick] (m) -- (pkeenc);
        \draw[->, thick] (m) -- (hashg);
        \draw[->, thick] (hashg) -- (mackey);
        \draw[->, thick] (pkeenc) -- (ct);
        \draw[->, thick] (mackey) -- (mac);
        \draw[->, thick] (ct) -- (mac);
        \draw[->, thick] (mac) -- (tag);
    \end{tikzpicture}

    \caption{Combining PKE with MAC using encrypt-then-MAC to encapsulate a shared secret. 
    The returned values are colored grey
    }\label{fig:etm-diagram}
\end{figure}

\textbf{Construction.} From a high level, the encrypt-then-MAC transformation takes as input a PKE and a MAC, then outputs a KEM. The output KEM encapsulates by encrypting a randomly sampled PKE plaintext and hashing the PKE plaintext-ciphertext pair into the shared secret. A symmetric key is derived from hashing the PKE plaintext, then used to produce a MAC tag on the PKE ciphertext. At decapsulation, the PKE secret key is used to recover the PKE plaintext, which is then hashed into a symmetric key that is used to verify the MAC tag against the input ciphertext. A summary of the data flow in the encrypt-then-MAC transformation can be found in Figure \ref{fig:etm-diagram}. A detailed description of individual routines can be found in Figure \ref{fig:etm-routines}.

\textbf{Security.} In Section \ref{sec:main-results}, we prove that if the input PKE is one-way secure against chosen-plaintext attack (OW-PCA) and the MAC is one-time existentially unforgeable, then the output KEM is IND-CCA secure. The IND-CCA security of the output KEM reduces tightly to the OW-PCA security of the input PKE, and non-tightly to the unforgeability of the MAC, though the non-tightness will likely not be a weakness because the construction makes it easy to plug in MAC of any desired security level.

\textbf{Implementation and performance comparison.} In Section \ref{sec:implementation}, we instantiate an instance of encrypt-then-MAC KEM using the PKE subroutines defined in ML-KEM \cite{FIPS203}. For clarity's sake we call this instantiation $\mlkemplus$. Compared with the full KEM routine in ML-KEM, which is based on some variant of the modular Fujisaki-Okamoto KEM transformation, $\mlkemplus$ demonstrates enormous runtime efficiency gain in decapsulation with only minimal performance penalty in encapsulation runtime and ciphertext size. On an AMD EPYC 9R14 CPU, $\mlkemplus$ (combined with Poly1305) achieves between 72\%-80\% reduction of CPU cycles needed for decapsulation while only incurring 2\%-7\% increase of CPU cycles needed for encapsulation. See Table \ref{tbl:cpu-cycles-summary} for a summary of performance comparison of individual routines against ML-KEM.

\begin{table}[H]
    \centering
    \footnotesize

    \begin{tabular}{|p{1.2cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
        \hline
        & $\text{ML-KEM}^+$ \newline 512 
        & $\text{ML-KEM}^+$ \newline 768 
        & $\text{ML-KEM}^+$ \newline 1024 
        \\
        \hline
        Encap \newline (ccl/tick) 
        & 93157 \newline (+1.8\%) 
        & 146405 \newline (+7.3\%) 
        & 205763 \newline (+3.3\%) 
        \\
        \hline
        Decap \newline (ccl/tick) 
        & 33733 \newline (-72.2\%) 
        & 43315 \newline (-76.8\%) 
        & 51375 \newline (-79.1\%) 
        \\
        \hline
        CT size \newline (bytes) 
        & 784 \newline (+2.1\%) 
        & 1104 \newline (+1.5\%) 
        & 1584 \newline (+1.0\%) 
        \\
        \hline
    \end{tabular}

    \caption{$\text{ML-KEM}^+$ achieves large performance improvements in decapsulation at the cost of minimal penalty in encapsulation performance and ciphertext size}\label{tbl:cpu-cycles-summary}
\end{table}

Since the decapsulation efficiency comes at the cost of increasing encryption runtime and ciphertext size, we also measured the round trip time of key exchange protocols with various modes of authentication. When compared to ML-KEM, $\mlkemplus$ achieves 24\%-28\% reduction of round trip time in unauthenticated key exchange (KE), 29\%-35\% in unilaterally authenticated key exchange (UAKE), and 35\%-48\% reduction in mutually authenticated key exchange (AKE). See Table \ref{tbl:rtt-summary} for a summary of round trip times.

\begin{table}[H]
    \centering
    \footnotesize

    \begin{tabular}{|p{1.6cm}|p{1.49cm}|p{1.49cm}|p{1.49cm}|}
        \hline
        &  $\text{ML-KEM}^+$ \newline 512 
        &  $\text{ML-KEM}^+$ \newline 768 
        &  $\text{ML-KEM}^+$ \newline 1024 
        \\
        \hline
        KE RTT \newline ($\us$) 
        &  70 \newline (-23.9\%) 
        &  99 \newline (-26.7\%) 
        &  138 \newline (-28.5\%) 
        \\
        \hline
        UAKE RTT \newline ($\us$) 
        &  103 \newline (-29.0\%) 
        &  144 \newline (-33.0\%) 
        &  202 \newline (-34.8\%) 
        \\
        \hline
        AKE RTT \newline ($\us$) 
        &  133 \newline (-39.5\%) 
        &  190 \newline (-35.4\%) 
        &  266 \newline (-48.0\%) 
        \\
        \hline
    \end{tabular}

    \caption{$\mlkemplus$ achieves substantial RTT savings despite increased encapsulation cost and ciphertext size}\label{tbl:rtt-summary}
\end{table}

\subsection{Related works}
\textbf{Optimal Asymmetric Encryption Padding (OAEP)} \cite{DBLP:conf/eurocrypt/BellareR94}\cite{DBLP:conf/crypto/BellareDPR98} is a generic chosen-ciphertext secure PKE. Under the random oracle model, the chosen-ciphertext security of the OAEP encryption scheme reduces to the one-wayness of the input trapdoor permutation. Although it was discovered that there exist trapdoor permutations with which the OAEP encryption scheme does not achieve full IND-CCA security \cite{DBLP:journals/joc/Shoup02}, Fujisaki et al. later proved that the OAEP is IND-CCA secure when combined with the RSA trapdoor permutation \cite{DBLP:conf/crypto/FujisakiOPS01}\cite{DBLP:journals/cacm/RivestSA78}. RSA-OAEP was standardized in PKCS\#1 v2 \cite{rfc8017} and is currently the most recommended of all RSA-based encryption schemes. Unfortunately, OAEP's requirement for a trapdoor permutation is immensely difficult to satisfy, and no other practical instantiation saw widespread adoption to this day.

The \textbf{Fujisaki-Okamoto transformation} \cite{DBLP:conf/crypto/FujisakiO99}\cite{DBLP:journals/joc/FujisakiO13} is another generic chosen-ciphertext secure transformation. While Fujisaki and Okamoto originally proposed a hybrid public-key encryption scheme whose IND-CCA security reduces non-tightly to the OW-CPA security of the underlying PKE and the IND-CPA security of the symmetric encryption scheme. Later works \cite{DBLP:conf/ima/Dent03}\cite{DBLP:conf/tcc/HofheinzHK17}\cite{DBLP:conf/eurocrypt/DworkNR04}\cite{DBLP:conf/asiacrypt/HovelmannsHM22}\cite{bernstein2018towards}  tightened the security reduction, accounted for imperfect correctness, adapted the original proposal to build a KEM, and proved its security in the quantum random oracle model (QROM).

The modular Fujisaki-Okamoto KEM transformation is remarkably successful because of the simplicity of its construction, the tightness of the security bound, and the proven (though non-tight) security against quantum adversaries. It was adopted by many submissions to NIST's post-quantum cryptography competition, including Kyber \cite{DBLP:conf/eurosp/BosDKLLSSSS18}, Saber \cite{DBLP:conf/africacrypt/DAnversKRV18}, FrodoKEM \cite{DBLP:conf/ccs/BosCDMNNRS16}, and classic McEliece \cite{classicmceliecespec} among others.

However, the Fujisaki-Okamoto transformation is not perfect. Among its shortfalls are two problems caused by its use of \emph{de-randomization} and \emph{re-encryption} for achieving chosen-ciphertext security. \begin{itemize}
    \item \textbf{Computational inefficiency}. The decapsulation routine needs to re-encrypt the decryption to ensure ciphertext has not been tempered with. For input PKE whose encryption routine carries substantial computational cost, such as most lattice-based cryptosystems, re-encryption slows down decapsulation significantly.
    \item \textbf{Side-channel vulnerability}. Re-encryption also introduces side-channels that can leak information about the decrypted PKE plaintext. As demonstrated in \cite{ueno2022curse}\cite{ravi2019generic}\cite{DBLP:journals/tches/TanakaUXITH23}, these side-channels can be converted into efficient plaintext-checking attacks that can fully recover the secret key in fewer than 1000 traces.
\end{itemize}

\section{Preliminaries}

\subsection{Public-key encryption scheme}
\textbf{Syntax.} A public-key encryption scheme $\pke(\keygen, \encrypt, \decrypt)$ is a collection of three routines defined over some plaintext space $\mathcal{M}$ and some ciphertext space $\mathcal{C}$. $(\pk, \sk) \leftsample \keygen()$ is a randomized routine that returns a keypair. The encryption routine $\encrypt: (\pk, m) \mapsto c$ encrypts the input plaintext under the input public key. The decryption routine $\decrypt: (\sk, c) \mapsto m$ decrypts the input ciphertext under the input secret key. Where the encryption routine is randomized, we denote the randomness by $r \in \mathcal{R}$, where $\mathcal{R}$ is called the coin space. The decryption routine is assumed to always be deterministic. Some decryption routines can detect malformed ciphertext and output the rejection symbol $\bot$ accordingly.

\textbf{Correctness.} Following the definition in \cite{DBLP:conf/eurocrypt/DworkNR04}, a $\pke$ is $\delta$-correct if:

\begin{equation*}
    E\left[\max_{m \in \mathcal{M}} P\left[\decrypt(\sk, c) \neq m \mid c \leftsample \encrypt(\pk, m)\right]\right] \leq \delta
\end{equation*}

Where the expectation is taken with respect to the probability distribution of all possible keypairs $(\pk, \sk) \leftsample \texttt{PKE.KeyGen()}$. For many lattice-based cryptosystems, including ML-KEM, decryption failures could leak information about the secret key, although the probability of a decryption failure is low enough that classical adversaries cannot exploit decryption failure more than they can defeat the underlying lattice problem. On the other hand, a quantum adversary may be able to exploit decryption failure in reasonable runtime by efficiently searching through all possible inputs using Grover's search algorithm. For that, ML-KEM made slight modifications in its KEM construction to prevent quantum adversary from precomputing large lookup table. We refer readers to \cite{avanzi2019crystals} and \cite{DBLP:conf/eurosp/BosDKLLSSSS18} for details of the mitigation techniques, though decryption failure is not a focus of this paper.

\textbf{Security.} The security of public-key encryption is conventionally discussed within the context of adversarial games played between a challenger and an adversary \cite{DBLP:conf/stoc/GoldwasserM82}. Each game is defined by the goal and the capabilities of the adversary.

\begin{figure}[H]
    \centering

    \begin{minipage}[t]{0.4\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{OW-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $m^\ast \leftsample \mathcal{M}$
                \State $c^\ast \leftsample \encrypt(\pk, m^\ast)$
                \State $\hat{m} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hspace{0.05\textwidth}
    \begin{minipage}[t]{0.4\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-ATK} Game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $(m_0, m_1) \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk)$
                \State $b \leftsample \{0,1\}$
                \State $c^\ast \leftsample \encrypt(\pk, m_b)$
                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(1^\lambda, \pk, c^\ast)$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The one-way game, indistinguishability game, plaintext-checking oracle (PCO), and decryption oracle. $\texttt{ATK} \in \{\texttt{CPA}, \texttt{PCA}, \texttt{CCA}\}$}
\end{figure}

Depending on the attack model, the adversary may have access to various oracles. In public-key cryptography, adversaries are always assumed to have the public key with which they can mount chosen-plaintext attack. If the adversary has access to a plaintext-checking oracle $\pco$ then it can mount plaintext-checking attack (PCA). Where the adversary has access to a decryption oracle, it can mount chosen-ciphertext attacks (CCA). A PKE is \texttt{OW-ATK}/\texttt{IND-ATK} secure if no efficient adversary with access to the corresponding oracle(s) can win the corresponding game with non-negligible advantage.

\subsection{Key encapsulation mechanism (KEM)}
\textbf{Syntax} A key encapsulation mechanism $\kem(\keygen, \encap, \decap)$ is a collection of three routines defined over some ciphertext space $\mathcal{C}$ and some key space $\mathcal{K}$. The key generation routine takes the security parameter $1^\lambda$ and outputs a keypair $(\pk, \sk) \leftsample \keygen(1^\lambda)$. $\encap(\pk)$ is a probabilistic routine that takes a public key $\pk$ and outputs a pair of values $(c, K)$ where $c \in \mathcal{C}$ is the ciphertext (also called encapsulation) and $K \in \mathcal{K}$ is the shared secret (also called session key). $\decap(\sk, c)$ is a deterministic routine that takes the secret key $\sk$ and the encapsulation $c$ and returns the shared secret $K$ if the ciphertext is valid. Some KEM constructions use explicit rejection, where if $c$ is invalid then $\decap$ will return a rejection symbol $\bot$; other KEM constructions use implicit rejection, where if $c$ is invalid then $\decap$ will return a fake session key that depends on the ciphertext and some other secret values.

\textbf{Security} The security of KEMs is similarly discussed in adversarial games (figure \ref{fig:kem-game}), although the win conditions differ slightly from the win conditions of a PKE's indistinguishability game. In a KEM's indistinguishability game, an adversary is given the public key and a challenge ciphertext, then asked to distinguish a pseudorandom shared secret $K_0$ associated with the challenge ciphertext from a truly random bit string of equal length.

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-ATK} game}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \keygen(1^\lambda)$
                \State $(c^\ast, K_0) \leftsample \encap(\pk)$
                \State $K_1 \leftsample \mathcal{K}$
                \State $b \leftsample \{0, 1\}$
                \State $\hat{b} \leftsample A^{\mathcal{O}_\texttt{ATK}}(
                    1^\lambda, \pk, c^\ast, K_b
                )$
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
        \caption*{$\mathcal{O}_\decap(c)$}
        \begin{algorithmic}[1]
            \State \Return $\decap(\sk, c)$
        \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{\texttt{IND-ATK} game for KEM and decapsulation oracle $\mathcal{O}_\decap$}\label{fig:kem-game}
\end{figure}

The decapsulation oracle $\mathcal{O}^\decap$ takes a ciphertext $c$ and returns the output of the $\decap$ routine using the secret key. The advantage $\epsilon_\texttt{IND-CCA}$ of an IND-CCA adversary $\mathcal{A}_\texttt{IND-CCA}$ is defined by

\begin{equation*}
    \adv_\texttt{IND-CCA}(A) = \norm{
        P[A^{\mathcal{O}_\decap}(a^\lambda, \pk, c^\ast, K_b) = b] - \frac{1}{2}
    }
\end{equation*}

\subsection{Message authentication code (MAC)}\label{sec:message-authentication-code}
A message authentication code $\mac$ is a collection of routines $(\sign, \verify)$ defined over some key space $\mathcal{K}$, some message space $\mathcal{M}$, and some tag space $\mathcal{T}$. The signing routine $\sign(k, m)$ takes the secret key $k \in \mathcal{K}$ and some message, and outputs a tag $t$. The verification routine $\verify(k, m, t)$ takes the triplet of secret key, message, and tag, and outputs $\texttt{1}$ if the message-tag pair is valid under the secret key, or $\texttt{0}$ otherwise. Many MAC constructions are deterministic. For these constructions it is simpler to denote the signing routine by $t \leftarrow \mac(k, m)$ and perform verification using a simple comparison.

The security of a MAC is defined in an adversarial game in which an adversary, with access to some signing oracle $\mathcal{O}_\sign(m)$, tries to forge a new valid message-tag pair that has never been queried before. The existential unforgeability under chosen message attack (EUF-CMA) game is shown below:

\begin{figure}[H]
    \centering
    \begin{minipage}{0.7\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{EUF-CMA} game}

            \begin{algorithmic}[1]
                \State $k^\ast \leftsample \mathcal{K}$
                \State $(\hat{m}, \hat{t}) \leftsample \mathcal{A}^{\mathcal{O}_\sign}()$
                \State \Return $
                    \llbrack \verify(k^\ast, \hat{m}, \hat{t}) 
                    \land (\hat{m}, \hat{t}) \not\in \mathcal{O}_\sign
                    \rrbrack
                $
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{The existential forgery game}\label{fig:euf-cma-game}
\end{figure}

The advantage $\adv_\texttt{EUF-CMA}$ of the existential forgery adversary is the probability that it wins the EUF-CMA game.

\section{The encrypt-then-MAC transformation}\label{sec:main-results}
Let $\mathcal{B}^\ast$ denote the set of finite bit strings. Let $\pke(\keygen, \encrypt, \decrypt)$ be a public-key encryption scheme defined over message space $\mathcal{M}$ and ciphertext space $\mathcal{C}$. Let $\mac: \mathcal{K}_\mac \times \mathcal{B}^\ast \rightarrow \mathcal{T}$ be a deterministic message authentication code that takes a key $k \in \mathcal{K}_\mac$, some message $m \in \mathcal{B}^\ast$, and outputs a digest $t \in \mathcal{T}$. Let $G: \mathcal{M} \rightarrow \mathcal{K}_\mac$ be a hash function that maps from $\pke$'s plaintext space to $\mac$'s key space. Let $H: \mathcal{B}^\ast \rightarrow \mathcal{K}_\kem$ be a hash function that maps bit strings into the set of possible shared secrets. The encrypt-then-MAC transformation $\etm[\pke, \mac, G, H]$ constructs a key encapsulation mechanism $\kem_\etm(\keygen_\kem, \encap, \decap)$, whose routines are described in figure \ref{fig:etm-routines}.

\begin{figure}[H]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\keygen()$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk^\prime) \leftsample \pke\texttt{.}\keygen()$
                \State $z \leftsample \mathcal{M}$
                \State $\sk \leftarrow \sk^\prime \Vert z$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\encap(\pk)$}
            \begin{algorithmic}[1]
                \State $m \leftsample \mathcal{M}$
                \State $k \leftarrow G(m)$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow H(m, c^\prime)$
                \State $c \leftarrow c^\prime \Vert t$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\kem_\etm.\decap(\sk, c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $(\sk^\prime, z) \leftarrow \sk$
                \State $\hat{m} \leftarrow \pke\texttt{.}\decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) \neq t$}
                    \State $K \leftarrow H(z, c^\prime)$
                \Else
                    \State $K \leftarrow H(\hat{m}, c^\prime)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{$\kem_\etm$ routines}\label{fig:etm-routines} 
\end{figure}

The key generation routine of $\kem_\etm$ is largely identical to that of the $\pke$, only a secret value $z$ is sampled as the implicit rejection symbol. In the encapsulation routine, a MAC key is derived from the randomly sampled plaintext $k \leftarrow G(m)$, then used to sign the unauthenticated ciphertext $c^\prime$. Because the encryption routine might be randomized, the session key is derived from both the message and the ciphertext. Finally, the unauthenticated ciphertext $c^\prime$ and the tag $t$ combine into the authenticated ciphertext $c$ that would be transmitted to the peer. In the decapsulation routine, the decryption $\hat{m}$ of the unauthenticated ciphertext is used to re-derive the MAC key $\hat{k}$, which is then used to re-compute the tag $\hat{t}$. The ciphertext is considered valid if and only if the recomputed tag is identical to the input tag.

For an adversary $A$ to produce a valid tag $t$ for some unauthenticated ciphertext $c^\prime$ under the symmetric key $k \leftarrow G(\decrypt(\sk^\prime, c^\prime))$ implies that $A$ must either know the symmetric key $k$ or produce a forgery. Under the random oracle model, $A$ also cannot know $k$ without knowing its preimage $\decrypt(\sk^\prime, c^\prime)$, so $A$ must either have produced $c^\prime$ honestly, or have broken the one-way security of $\pke$. This means that the decapsulation oracle will not give out information on decryptions that the adversary does not already know.

\begin{figure}[H]
    \centering
    \begin{minipage}{0.6\textwidth}
    \begin{algorithm}[H]
        \caption*{$\pco(m, c)$}
        \begin{algorithmic}[1]
            \State $k \leftarrow G(m)$
            \State $t \leftarrow \mac(k, c)$
            \State \Return $\llbrack \mathcal{O}^\decap((c, t)) = H(m, c)\rrbrack$
        \end{algorithmic}
    \end{algorithm}
    \end{minipage}
    \caption{Every decapsulation oracle can be converted into a plaintext-checking oracle}\label{fig:pco-using-decap-oracle}
\end{figure}

However, a decapsulation oracle can still give out some information: for a known plaintext $m$, all possible encryptions $c^\prime \leftsample \encrypt(\pk, m)$ can be correctly signed, while ciphertexts that don't decrypt back to $m$ cannot be correctly signed. This means that a decapsulation oracle can be converted into a plaintext-checking oracle (see figure \ref{fig:pco-using-decap-oracle}), so every chosen-ciphertext attack against the KEM can be converted into a plaintext-checking attack against the underlying PKE.

On the other hand, if the underlying PKE is one-way secure against plaintext-checking attack that makes $q$ plaintext-checking queries, then encrypt-then-MAC KEM is semantically secure under chosen ciphertext attacks making the same number of decapsulation queries:

\begin{theorem}\label{thm:ow-pca-implies-kem-ind-cca2}
    For every \texttt{IND-CCA} adversary $A$ against $\kem_\etm$ that makes $q$ decapsulation queries, there exists an \texttt{OW-PCA} adversary $B$ who makes at least $q$ plaintext-checking queries against the underlying $\pke$, and an one-time existential forgery adversary $C$ against the underlying $\mac$ such that

    \begin{equation*}
        \texttt{Adv}_\texttt{IND-CCA}(A) \leq q \cdot \adv_\texttt{OT-MAC}(C) + 2 \cdot \texttt{Adv}_\texttt{OW-PCA}(B)
    \end{equation*}
\end{theorem}

Theorem \ref{thm:ow-pca-implies-kem-ind-cca2} naturally flows into an equivalence relationship between the security of the KEM and the security of the PKE:

\begin{lemma}
    $\kem_\etm$ is IND-CCA secure if and only if the input $\pke$ is OW-PCA secure
\end{lemma}

\subsection{Proof of theorem \ref{thm:ow-pca-implies-kem-ind-cca2}}
We will prove theorem \ref{thm:ow-pca-implies-kem-ind-cca2} using a sequence of game. A summary of the the sequence of games can be found in figure \ref{fig:etm-ind-cc2-sequence-of-games} and \ref{fig:ow-pca-simulates-game-3}. From a high level we made three incremental modifications to the IND-CCA game for $\kem_\etm$: replace true decapsulation with simulated decapsulation, replace the pseudorandom MAC key $k^\ast \leftarrow G(m^\ast)$ with a truly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$, and finally replace pseudorandom shared secret $K_0 \leftarrow H(m^\ast, c^\prime)$ with a truly random shared secret $K_0 \leftsample \mathcal{K}_\kem$. A OW-PCA adversary can then simulate the modified IND-CCA game for the KEM adversary, and the advantage of the OW-PCA adversary is associated with the probability of certain behaviors of the KEM adversary.

\begin{figure}[h]
    \centering
    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{IND-CCA} game for $\kem_\etm$}
            \begin{algorithmic}[1]
                \State $(\pk, \sk) \leftsample \kem_\etm\texttt{.}\keygen()$
                \State $m^\ast \leftsample \mathcal{M}$
                \State $c^\prime \leftsample \pke\texttt{.}\encrypt(\pk, m^\ast)$
                \State $k^\ast \leftarrow G(m^\ast)$
                    \Comment{Game 0-1}
                \State $k^\ast \leftsample \mathcal{K}_\mac$
                    \Comment{Game 2-3}
                \State $t \leftarrow \mac(k^\ast, c^\prime)$
                \State $c^\ast \leftarrow c^\prime \Vert t$
                \State $K_0 \leftarrow H(m^\ast, c^\prime)$
                    \Comment{Game 0-2}
                \State $K_0 \leftsample \mathcal{K}_\kem$
                    \Comment{Game 3}
                \State $K_1 \leftsample \mathcal{K}_\kem$
                \State $b \leftsample \{0,1\}$
                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap}(\pk, c^\ast, K_b)$
                    \Comment{Game 0}
                \State $\hat{b} \leftarrow A^{\mathcal{O}^\decap_1}(\pk, c^\ast, K_b)$
                    \Comment{Game 1-3}
                \State \Return $\llbrack \hat{b} = b \rrbrack$
            \end{algorithmic}
        \end{algorithm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^G(m)$}
            \begin{algorithmic}[1]
                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
                    \State \Return $\tilde{k}$
                \EndIf
                \State $k \leftsample \mathcal{K}_\mac$
                \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
                \State \Return $k$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^\decap(c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \State $\hat{m} = \decrypt(\sk^\prime, c^\prime)$
                \State $\hat{k} \leftarrow G(\hat{m})$
                \If{$\mac(\hat{k}, c^\prime) = t$}
                    \State $K \leftarrow H(\hat{m}, c^\prime)$
                \Else 
                    \State $K \leftarrow H(z, c^\prime)$
                \EndIf 
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
        \vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^\decap_1(c)$}
            \begin{algorithmic}[1]
                \State $(c^\prime, t) \leftarrow c$
                \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
                    \tilde{m} = \decrypt(\sk^\prime, c^\prime)
                    \land \mac(\tilde{k}, c^\prime) = t
                $}
                    \State $K \leftarrow H(\tilde{m}, c^\prime)$
                \Else
                    \State $K \leftarrow H(z, c^\prime)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
        \vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\mathcal{O}^H(m, c)$}
            \begin{algorithmic}[1]
                \If{$\exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H : \tilde{m} = m \land \tilde{c} = c$}
                    \State \Return $\tilde{K}$
                \EndIf
                \State $K \leftsample \mathcal{K}_\kem$
                \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \caption{Sequence of games}\label{fig:etm-ind-cc2-sequence-of-games}
\end{figure}

\begin{proof}
    \emph{Game 0} is the standard IND-CCA game for KEMs. The decapsulation oracle $\mathcal{O}^\decap$ executes the decapsulation routine using the challenge keypair and return the results faithfully. The queries made to the hash oracles $\mathcal{O}^G, \mathcal{O}^H$ are recorded to their respective tapes $\mathcal{L}^G, \mathcal{L}^H$.

    \emph{Game 1} is identical to game 0 except that the true decapsulation oracle $\mathcal{O}^\decap$ is replaced with a simulated oracle $\mathcal{O}^\decap_1$. Instead of directly decrypting $c^\prime$ as in the decapsulation routine, the simulated oracle searches through the tape $\mathcal{L}^G$ to find a matching query $(\tilde{m}, \tilde{k})$ such that $\tilde{m}$ is the decryption of $c^\prime$. The simulated oracle then uses $\tilde{k}$ to validate the tag $t$ against $c^\prime$.

    If the simulated oracle accepts the queried ciphertext as valid, then there is a matching query that also validates the tag, which means that the queried ciphertext is honestly generated. Therefore, the true oracle must also accept the queried ciphertext. On the other hand, if the true oracle rejects the queried ciphertext (and output the implicit rejection $H(z, c^\prime)$), then the tag is simply invalid under the MAC key $k = G(\decrypt(\sk^\prime, c^\prime))$. Therefore, there could not have been a matching query that also validates the tag, and the simulated oracle must also rejects the queried ciphertext.

    This means that from the adversary $A$'s perspective, game 1 and game 0 differ only when the true oracle accepts while the simulated oracle rejects, which means that $t$ is a valid tag for $c^\prime$ under $k = G(\decrypt(\sk^\prime, c^\prime))$, but $k$ has never been queried. Under the random oracle model, such $k$ is a uniformly random sample of $\mathcal{K}_\mac$ that the adversary does not know, so for $A$ to produce a valid tag is to produce a forgery against the $\mac$ under an unknown and uniformly random key. Furthermore, the security game does not include a signing oracle, so this is a zero-time forgery. While zero-time forgery is not a standard security definition for a MAC, we can bound it by the advantage of a one-time forgery adversary $C$:

    \begin{equation*}
        P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]
        \leq \adv_\texttt{OT-MAC}(C)
    \end{equation*}

    Across all $q$ decapsulation queries, the probability that at least one query is a forgery is thus at most $q \cdot P\left[\mathcal{O}^\decap(c) \neq \mathcal{O}^\decap_1(c)\right]$. By the difference lemma:

    \begin{equation*}
        \adv_{G_0}(A) - \adv_{G_1}(A) \leq q\cdot  \adv_\texttt{OT-MAC}(C)
    \end{equation*}

    \emph{Game 2} is identical to game 1, except that the challenger samples a uniformly random MAC key $k^\ast \leftsample \mathcal{K}_\mac$ instead of deriving it from $m^\ast$. From $A$'s perspective the two games are indistinguishable, unless $A$ queries $G$ with the value of $m^\ast$. Denote the probability that $A$ queries $G$ with $m^\ast$ by $P[\texttt{QUERY G}]$, then:

    \begin{equation*}
        \adv_{G_1}(A) - \adv_{G_2}(A) \leq P\left[\texttt{QUERY G}\right]
    \end{equation*}

    \emph{Game 3} is identical to game 2, except that the challenger samples a uniformly random shared secret $K_0 \leftsample \mathcal{K}_\kem$ instead of deriving it from $m^\ast$ and $c^\prime$. From $A$'s perspective the two games are indistinguishable, unless $A$ queries $H$ with $(m^\ast, \cdot)$. Denote the probability that $A$ queries $H$ with $(m^\ast, \cdot)$ by $P[\texttt{QUERY H}]$, then:

    \begin{equation*}
        \adv_{G_2}(A) - \adv_{G_3}(A) \leq P\left[\texttt{QUERY H}\right]
    \end{equation*}

    Since in game 3, both $K_0$ and $K_1$ are uniformly random and independent of all other variables, no adversary can have any advantage: $\adv_{G_3}(A) = 0$.

    \begin{figure}[h]
        \begin{minipage}[t]{0.49\textwidth}
            \begin{algorithm}[H]
                \caption*{$B(\pk, {c^\prime}^\ast)$}
                \begin{algorithmic}[1]
                    \State $z \leftsample \mathcal{M}$
                    \State $k \leftsample \mathcal{K}_\mac$
                    \State $t \leftarrow \mac(k, {c^\prime}^\ast)$
                    \State $c^\ast \leftarrow ({c^\prime}^\ast, t)$
                    \State $K \leftsample \mathcal{K}_\kem$
                    \State $\hat{b} \leftarrow A^{
                        \mathcal{O}^\decap_B, \mathcal{O}^G_B, \mathcal{O}^H_B
                    }(\pk, c^\ast, K)$
                    \If{$\texttt{ABORT}(m)$}
                        \State \Return $m$
                    \EndIf
                \end{algorithmic}
            \end{algorithm}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^H_B(m, c)$}
                \begin{algorithmic}
                    \If{$\pco(m, {c^\prime}^\ast) = 1$}
                        \State $\texttt{ABORT}(m)$
                    \EndIf
                    \If{$
                        \exists (\tilde{m}, \tilde{c}, \tilde{K}) \in \mathcal{L}^H 
                        : \tilde{m} = m \land \tilde{c} = c
                    $}
                        \State \Return $\tilde{K}$
                    \EndIf
                    \State $K \leftsample \mathcal{K}_\kem$
                    \State $\mathcal{L}^H \leftarrow \mathcal{L}^H \cup \{(m, c, K)\}$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}\hfill
        \begin{minipage}[t]{0.49\textwidth}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^\decap_B(c)$}
                \begin{algorithmic}[1]
                    \State $(c^\prime, t) \leftarrow c$
                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : 
                        % \tilde{m} = \decrypt(\sk^\prime, c^\prime)
                        \pco(c^\prime, \tilde{m}) = 1
                        \land \mac(\tilde{k}, c^\prime) = t
                    $}
                        \State $K \leftarrow H(\tilde{m}, c^\prime)$
                    \Else
                        \State $K \leftarrow H(z, c^\prime)$
                    \EndIf
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^G_B(m)$}
                \begin{algorithmic}[1]
                    \If{$\pco(m, {c^\prime}^\ast) = 1$}
                        \State $\texttt{ABORT}(m)$
                    \EndIf
                    \If{$\exists (\tilde{m}, \tilde{k}) \in \mathcal{L}^G : \tilde{m} = m$}
                        \State \Return $\tilde{k}$
                    \EndIf
                    \State $k \leftsample \mathcal{K}_\mac$
                    \State $\mathcal{L}^G \leftarrow \mathcal{L}^G \cup \{(m, k)\}$
                    \State \Return $k$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \caption{OW-PCA adversary $B$ simulates game 3 for IND-CCA adversary $A$}\label{fig:ow-pca-simulates-game-3}
    \end{figure}

    We will bound $P[\texttt{QUERY G}]$ and $P[\texttt{QUERY H}]$ by constructing a OW-PCA adversary $B$ against the underlying PKE that uses $A$ as a sub-routine. $B$'s behaviors are summarized in figure \ref{fig:ow-pca-simulates-game-3}.

    $B$ simulates game 3 for $A$: receiving the public key $\pk$ and challenge encryption ${c^\prime}^\ast$, $B$ samples random MAC key and session key to produce the challenge encapsulation, then feeds it to $A$. When simulating the decapsulation oracle, $B$ uses the plaintext-checking oracle to look for matching queries in $\mathcal{L}^G$. When simulating the hash oracles, $B$ uses the plaintext-checking oracle to detect when $m^\ast = \decrypt(\sk^\prime, {c^\prime}^\star)$ has been queried. When $m^\ast$ is queried, $B$ terminates $A$ and returns $m^\ast$ to win the OW-PCA game. In other words:

    \begin{equation*}
        \begin{aligned}
            P\left[\texttt{QUERY G}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
            P\left[\texttt{QUERY H}\right] &\leq \adv_\texttt{OW-PCA}(B) \\
        \end{aligned}
    \end{equation*}

    Combining all equations above produce the desired security bound.
\end{proof}

\subsection{ElGamal is OW-PCA secure}
We show that the DHAES/DHIES hybrid encryption scheme is a special case of the encrypt-then-MAC transformation. Specifically, we will sketch a proof of the following lemma:

\begin{lemma}\label{lemma:elgamal-is-ow-pca}
    For every OW-PCA adversary $A$ against the ElGamal cryptosystem, there exists a Gap Diffie-Hellman problem solver $B$ such that:

    \begin{equation*}
        \text{Adv}_\text{GapDH}(B) = \text{Adv}_\text{OW-PCA}(A)
    \end{equation*}

    In other words, ElGamal is OW-PCA secure under the Gap Diffie-Hellman assumption.
\end{lemma}

Each ElGamal cryptosystem \cite{DBLP:journals/tit/Elgamal85} is parameterized by a cyclic group $G = \langle g \rangle$ of prime order $q > 2$. A summary of the routine is shown below:

\begin{figure}[H]
    \centering

    \begin{minipage}[t]{0.32\textwidth}
        \begin{algorithm}[H]
            \caption*{\texttt{KeyGen()}}
            \begin{algorithmic}[1]
                \State $x \leftsample \mathbb{Z}_q$
                \State $\sk \leftarrow x$
                \State $\pk \leftarrow g^x$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{Enc}(\pk = g^x, m \in G)$}
            \begin{algorithmic}[1]
                \Require $m \in G$
                \State $y \leftsample \mathbb{Z}_q$
                \State $w \leftarrow g^y$
                \State $v \leftarrow m \cdot (g^x)^y$
                \State \Return $c = (w, v)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.33\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{Dec}(\sk = x, c = (w,v) \in G^2)$}
            \begin{algorithmic}[1]
                \State $\hat{m} \leftarrow (w^x)^{-1}\cdot v$
                \State \Return $\hat{m}$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{ElGamal cryptosystem}
\end{figure}

The security of ElGamal cryptosystem reduces to the conjectured intractability of the computational Diffie-Hellman problem and the decisional Diffie-Hellman problem:

\begin{definition}[\textbf{computational Diffie-Hellman problem (CDH)}]
    Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $(g, g^x, g^y)$, compute $g^{xy}$.
\end{definition}
\begin{definition}[\textbf{decisional Diffie-Hellman problem (DDH)}]
    Let $x, y, z \leftsample \mathbb{Z}_q$ be uniformly random samples. Let $h \leftsample \{g^z, g^{xy}\}$ be randomly chosen between $g^z$ and $g^{xy}$. Given $(g, g^x, g^y, h)$, determine whether $h$ is $g^{xy}$ or $g^z$
\end{definition}

It is also conjectured in \cite{DBLP:conf/ctrsa/AbdallaBR01} that for certain choice of cyclic group $G$, the computational Diffie-Hellman problem remains intractable even if the adversary as access to a restricted decisional Diffie-Hellman oracle. This assumption is captured in the Gap Diffie-Hellman problem:

\begin{definition}[\textbf{Gap Diffie-Hellman problem}]
    Let $G = \langle g \rangle$ be a cyclic group of prime order $q > 2$. Let $x, y \leftsample \mathbb{Z}_q$ be uniformly random samples. Given $(g, g^x, g^y)$ and a restricted DDH oracle $\mathcal{O}^\text{DDH}: (u, v) \mapsto \llbrack u^x = v \rrbrack$, compute $g^{xy}$.
\end{definition}

We now present the proof for lemma \ref{lemma:elgamal-is-ow-pca}
\begin{proof}
    We will prove by sequence of game. A summary can be found in figure \ref{fig:elgamal-pca-games}

    \begin{figure}
        \centering

        \begin{minipage}[t]{0.4\textwidth}
            \begin{algorithm}[H]
                \caption*{$G_0 - G_2$}
                \begin{algorithmic}[1]
                    \State $x \leftsample \mathbb{Z}_q$
                    \State $m^\ast \leftsample G$
                    \State $y \leftsample \mathbb{Z}_q, w \leftarrow g^y$
                    \State $v \leftarrow m^\ast \cdot (g^x)^y$
                        \Comment{$G_0$ - $G_1$}
                    \State $v \leftsample G$
                        \Comment{$G_2$}
                    \State $c^\ast \leftarrow (w, v)$
                    \State $\hat{m} \leftsample A^{\pco}(g^x, c^\ast)$
                        \Comment{$G_0$}
                    \State $\hat{m} \leftsample A^{\pco_1}(g^x, c^\ast)$
                        \Comment{$G_1$ - $G_2$}
                    \State \Return $\llbrack \hat{m} = m^\ast \rrbrack$
                        \Comment{$G_0$ - $G_1$}
                    \State \Return $\llbrack \hat{m} = w^{-x}\cdot v \rrbrack$
                        \Comment{$G_2$}
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \begin{minipage}[t]{0.4\textwidth}
            \begin{algorithm}[H]
                \caption*{$\pco(m, c=(w, v))$}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack m = (w^x)^{-1}\cdot v\rrbrack$
                \end{algorithmic}
            \end{algorithm}
            \begin{algorithm}[H]
                \caption*{$\pco_1(m, c=(w, v))$}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack (w^x) = m^{-1} \cdot v \rrbrack$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}

        \caption{Lemma \ref{lemma:elgamal-is-ow-pca} sequence of games}\label{fig:elgamal-pca-games}
    \end{figure}

    \emph{Game 0} is the OW-PCA game. Adversary $A$ has access to the plaintext-checking oracle $\pco$ and wins the game if it can correctly recover the challenge plaintext $m^\ast$.

    \emph{Game 1} is identical to game 0, except that the formulation of the $\pco$ is changed. When servicing the plaintext-checking query $(m, c = (w, v))$, $\pco_1$ checks whether $w^x$ is equal to $m^{-1} \cdot v$. Observe that in the cyclic group $G$, the algebraic expressions in $\pco$ and $\pco_1$ are equivalent, which means that $\pco_1$ behaves identically to $\pco$.

    \emph{Game 2} is identical to game 1 except for two modifications: first, when computing the challenge ciphertext, $v$ is no longer computed from $m^\ast$ but is randomly sampled; second, the win condition changed from $\hat{m} = m^\ast$ to $\hat{m} = w^{-x}\cdot v$. Notice that in game 0-1, $v \leftsample m^\ast \cdot (g^x)^y$ follows uniform random distribution on the cyclic group $G$ because $m^\ast$ is uniformly random in the cyclic group, so adversary $A$ retains its advantage in ``recovering the decryption'' when $v$ becomes truly uniformly random in the cyclic group. It is easy to verify that the two win conditions are equivalent. Up to this point, we have simply moved things around, and game 0 through game 2 are algebraically equivalent:

    \begin{equation*}
        \adv_0(A) = \adv_1(A) = \adv_2(A)
    \end{equation*}

    The Gap Diffie-Hellman adversary $B$ can perfectly simulate game 2 for $A$ (see figure \ref{fig:ow-pca-to-gap-dh}): $B$ receives as the Gap Diffie-Hellman problem inputs $g^x$ and $g^y$. $g^x$ simulates an ElGamal public key, where as $g^y$ simulates the first component of the challenge ciphertext. As in game 2, the second component of the challenge ciphertext can be randomly sampled. Finally, the $\pco_1$ from game 2 can be perfectly simulated using the restricted DDH oracle $\mathcal{O}^\text{DDH}$.

    \begin{figure}
        \centering

        \begin{minipage}[t]{0.45\textwidth}
            \begin{algorithm}[H]
                \caption*{$B^{\mathcal{O}^\text{DDH}}(g, g^x, g^y)$}
                \begin{algorithmic}[1]
                    \State $w \leftarrow g^y$
                    \State $v \leftsample G$
                    \State $c^\ast \leftarrow (w, v)$
                    \State $\hat{m} \leftsample A^{\pco_2}(g^x, c^\ast)$
                    \State \Return $\hat{m}^{-1}\cdot v$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}
        \begin{minipage}[t]{0.45\textwidth}
            \begin{algorithm}[H]
                \caption*{$\mathcal{O}^\text{DDH}(u, v)$}
                \begin{algorithmic}[1]
                    \State \Return $\llbrack u^x = v \rrbrack$
                \end{algorithmic}
            \end{algorithm}\vspace{-0.3cm}
            \begin{algorithm}[H]
                \caption*{$\pco_2(m, c=(w, v))$}
                \begin{algorithmic}[1]
                    \State \Return $\mathcal{O}^\text{DDH}(w, m^{-1}\cdot v)$
                \end{algorithmic}
            \end{algorithm}
        \end{minipage}

        \caption{Gap Diffie-Hellman adversary $B$ simulates game 2 for $A$}\label{fig:ow-pca-to-gap-dh}
    \end{figure}

    If $A$ wins game 2, then its output is $\hat{m} = w^{-x}\cdot v = g^{-xy}\cdot v$, so $m^{-1}\cdot v$ is $g^{xy}$, the correct answer to the Gap Diffie-Hellman problem. In other words, $A$ wins game 2 if and only if $B$ solves the Gap Diffie-Hellman problem:

    \begin{equation*}
        \adv_2(A) = \adv_\text{GapDH}(B)
    \end{equation*}
\end{proof}

\section{Implementation}\label{sec:implementation}
ML-KEM is an \texttt{IND-CCA} secure key encapsulation mechanism standardized by NIST in FIPS 203. The \texttt{IND-CCA} security of ML-KEM is achieved in two steps. First, ML-KEM constructs an \texttt{IND-CPA} secure public key encryption scheme $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ whose security is based on the conjectured intractability of the module learning with error (MLWE) problems against both classical and quantum adversaries. Then, the $U^{\not\bot}_m$ variant of the Fujisaki-Okamoto transformation is used to construct the KEM $\texttt{MLKEM}(\keygen, \encap, \decap)$ by calling $\texttt{K-PKE}(\keygen, \encrypt, \decrypt)$ as sub-routines. Because $\texttt{K-PKE}.\encrypt$ includes substantially more arithmetics than $\texttt{K-PKE}.\decrypt$, by using \textit{re-encryption} and \textit{de-randomization}, ML-KEM's decapsulation routine incurs significant computational cost.

We implemented the encrypt-then-MAC KEM construction using $\texttt{K-PKE}$ as the input PKE and compared its performance against ML-KEM under a variety of scenarios. The experimental data showed that while the encrypt-then-MAC construction adds a small amount of computational overhead to the encapsulation routine and a small increase in ciphertext size when compared with ML-KEM, it boasts enormous runtime savings in the decapsulation routine, which makes it particularly suitable for deployment in constrained environment.

We refer readers to \cite{FIPS203} for the details of the \texttt{K-PKE} routines. The encrypt-then-MAC KEM routines are described in figure \ref{fig:ml-kem-plus-routines} below.

\begin{figure}[H]
    \centering

    \begin{minipage}[t]{0.5\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\keygen\texttt{()}$}
            \begin{algorithmic}[1]
                \State $z \leftsample \{0,1\}^{256}$
                \State $(\pk, \sk^\prime) \leftsample \texttt{K-PKE}.\keygen()$
                \State $h \leftarrow H(\pk)$
                \State $\sk \leftarrow (sk^\prime \Vert \pk \Vert h \Vert z)$
                \State \Return $(\pk, \sk)$
            \end{algorithmic}
        \end{algorithm}\vspace{-0.5cm}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\encap\texttt{(\pk)}$}
            \begin{algorithmic}[1]
                \Require Public key $\pk$
                \State $m \leftsample \{0,1\}^{256}$
                \State $(\overline{K}, r, k) \leftarrow \texttt{XOF}(m \Vert H(\pk))$
                \State $c^\prime \leftarrow \texttt{K-PKE}.\encrypt(\pk, m, r)$
                \State $t \leftarrow \mac(k, c^\prime)$
                \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert c^\prime)$
                \State $c \leftarrow (c^\prime, t)$
                \State \Return $(c, K)$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[t]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{ML-KEM}^\texttt{+}\texttt{.}\decap\texttt{(\sk, c)}$}
            \begin{algorithmic}[1]
                \Require Secret key $\sk = (\sk^\prime \Vert \pk \Vert h \Vert z)$
                \Require Ciphertext $c = (c^\prime \Vert t)$
                \State $(\sk^\prime, \pk, h, z) \leftarrow \sk$
                \State $(c^\prime, t) \leftarrow c$
                \State $\hat{m} \leftarrow \texttt{K-PKE}.\decrypt(\sk^\prime, c^\prime)$
                \State $(\overline{K}, \hat{r}, \hat{k}) \leftarrow \texttt{XOF}(\hat{m} \Vert h)$
                \State $\hat{t} \leftarrow \mac(\hat{k}, c^\prime)$
                \If{$\hat{t} = t$}
                    \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert t)$
                \Else
                    \State $K \leftarrow \texttt{KDF}(z \Vert t)$
                \EndIf
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{$\texttt{ML-KEM}^+$ routines}\label{fig:ml-kem-plus-routines}
\end{figure}

Our implementation extended from the reference implementation by the PQCrystals team (https://github.com/pq-crystals/kyber). All C code is compiled with \texttt{GCC 11.4.1} and \texttt{OpenSSL 3.0.8}. All binaries are executed on an AWS c7a.medium instance with an AMD EPYC 9R14 CPU at 3.7 GHz and 1 GB of RAM.

\subsection{Choosing a message authenticator}\label{sec:choosing-a-message-authenticator}
For the $\texttt{ML-KEM}^+$ implementation, we instantiated $\mac$ with a selection that covered a wide range of MAC designs, including Poly1305 \cite{bernstein2005poly1305}, GMAC \cite{mcgrew2004galois}, CMAC \cite{iwata2003omac}\cite{black2000cbc}, and KMAC \cite{SP80053r4}.

Poly1305 and GMAC are both Carter-Wegman style authenticators \cite{wegman1981new} that compute the tag using finite field arithmetic. Generically speaking, Carter-Wegman MAC is parameterized by some finite field $\mathbb{F}$ and the maximal message length $L > 0$. Each symmetric key $k = (k_1, k_2) \leftsample \mathbb{F}^2$ is a pair of uniformly ranodm field elements, and the message is parsed into tuples of field elements up to length $L$: $m = (m_1, m_2, \ldots, m_l) \in \mathbb{F}^{\leq L}$. The tag $t$ is computed by evaluating a polynomial whose coefficients the message blocks and whose indeterminate is the key:

\begin{equation}\label{eq:carter-wegman-one-time-mac}
    \mac((k_1, k_2), m) = H_\text{xpoly}(k_1, m) + k_2
\end{equation}

Where $H_\text{xpoly}$ is given by:

\begin{equation*}
    H_\text{xpoly}(k_1, m) = k_1^{l+1} + k_1^l \cdot m_1 + k_1^{l-1} \cdot m_2 + \ldots + k_1 \cdot m_l
\end{equation*}

The authenticator formulated in equation \ref{eq:carter-wegman-one-time-mac} is a one-time MAC. To make the construction many-time secure, a non-repeating nonce $r$ and a PRF is needed:

\begin{equation*}
    \mac((k_1, k_2), m, r) = H_\text{xpoly}(k_1, m) \oplus \texttt{PRF}(k_2, r)
\end{equation*}

Specifically, Poly1035 operates in the prime field $\mathbb{F}_q$ where $q = 2^{130} - 5$ whereas GMAC operates in the binary field $\mathbb{F}_{2^{128}}$. In OpenSSL's implementation, standalone Poly1305 is a one-time secure MAC, whereas GMAC uses a nonce and AES as the PRF and is thus many-time secure (in OpenSSL GMAC is AES-256-GCM except all data is fed into the ``associated data'' section and thus not encrypted).

CMAC is based on the CBC-MAC with the block cipher instantiated from AES-256. To compute a CMAC tag, the message is first broke into 128-bit blocks with appropriate padding. Each block is first XOR'd with the previous block's output, then encrypted under AES using the symmetric key. The final output is XOR'd with a sub key derived from the symmetric key, before being encrypted for one last time. %A summary of the computation can be found in figure \ref{fig:aes-256-cmac}

% \begin{figure}[h]
%     \centering
%     \begin{minipage}[t]{0.49\textwidth}
%         \begin{algorithm}[H]
%             \caption*{Sub-key derivation}
%             \begin{algorithmic}[1]
%                 \Require 256-bit key $k$
%                 \Require $\texttt{const\_Rb = 0x87}$
%                 \State $l \leftarrow \texttt{AES-256}(k, 0^{128})$
%                 \If{$\texttt{MostSignificantBit}(l) = 0$}
%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1}$
%                 \Else 
%                     \State $k_1 \leftarrow \texttt{l {<}{<} 1} \oplus \texttt{const\_Rb}$
%                 \EndIf
%                 \If{$\texttt{MostSignificantBit}(k_1) = 0$}
%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1}$
%                 \Else 
%                     \State $k_2 \leftarrow k_1 \texttt{ {<}{<} 1} \oplus \texttt{const\_Rb}$
%                 \EndIf
%                 \State \Return $k_1, k_2$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{0.49\textwidth}
%         \begin{algorithm}[H]
%             \caption*{CMAC(k, m)}
%             \begin{algorithmic}[1]
%                 \Require 256-bit symmetric key $k$
%                 \State $(k_1, k_2) \leftarrow \texttt{deriveSubKey}(k)$
%                 \State $n \leftarrow \left\lceil \texttt{bytesLen}(m) / 16\right\rceil$
%                 \If{$n = 0$}
%                     \State $n \leftarrow 1$
%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
%                 \ElsIf{$\texttt{bytesLen(m)} \mod 16 = 0$}
%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_1$
%                 \Else 
%                     \State $m_\texttt{last} \leftarrow m_n \oplus k_2$
%                 \EndIf

%                 \State $x = 0^{128}$
%                 \For{$i \in \{1, 2, \ldots, n-1\}$}
%                     \State $y \leftarrow x \oplus m_i$
%                     \State $x \leftarrow \texttt{AES-256}(k, y)$
%                 \EndFor
%                 \State $y \leftarrow m_\texttt{last} \oplus x$
%                 \State $t \leftarrow \texttt{AES-256}(k, y)$
%                 \State \Return $t$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}
%     \caption{AES-256 CMAC}\label{fig:aes-256-cmac}
% \end{figure}

KMAC is defined in \cite{SP80053r4} to be based on the family of sponge functions with Keccak permutaiton as the underlying function. We chose KMAC-256, which uses Shake256 as the underlying extendable output functions. KMAC allows variable-length key and tag, but we chose the 256 bits for key length and 128 bits for tag size for consistency with other authenticators.

To isolate the performance characteristics of each authenticator in our instantiation of $\texttt{ML-KEM}^+$, we measured the CPU cycles needed for each authenticator to compute a tag on random inputs whose sizes correspond to the ciphertext sizes of \texttt{ML-KEM}. The measurements are summarized in table \ref{tbl:mac-performance}.

From our testing, we found Poly1305 to exhibit the best performance characteristics. However, there are additional security considerations that may require the use of other less efficient MAC instances. For example, it is possible for an adversary with large computing infrastructure or quantum computers to pre-compute a large lookup table mapping symmetric key to the source plaintext. Upon receiving a ciphertext, the adversary can then search through the lookup table for a matching key, which would've revealed the corresponding decryption. We partially mitigated such attack by deriving the symmetric key from both the public key and the plaintext, but in case of a long-term keypair, the adversary might still be able to compute a large lookup table AFTER obtaining the long-term public key. Further mitigation could include using larger-size keys, which can be accomplished either by using a Carter-Wegman MAC that operates on a larger finite field or using a MAC with a variable key-length such as KMAC.

\begin{table}[H]
    \centering
    \footnotesize
    \caption{CPU cycles needed to compute tag on various input sizes}\label{tbl:mac-performance}
    \begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\bf Input size: 768 bytes} \\
        \hline
        MAC & Median & Average \\
        \hline
        Poly1305 & 909 & 2823 \\
        \hline
        GMAC & 3899 & 4859 \\
        \hline
        CMAC & 6291 & 6373 \\
        \hline
        KMAC & 6373 & 7791 \\
        \hline
    \end{tabular}
    \begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\bf Input size: 1088 bytes} \\
        \hline
        MAC & Median & Average \\
        \hline
        Poly1305 & 961 & 2704 \\
        \hline
        GMAC & 3899 & 4827 \\
        \hline
        CMAC & 7305 & 7588 \\
        \hline
        KMAC & 9697 & 9928 \\
        \hline
    \end{tabular}
    \begin{tabular}{|c|c|c|}
        \hline
        \multicolumn{3}{|c|}{\bf Input size: 1568 bytes} \\
        \hline
        MAC & Median & Average \\
        \hline
        Poly1305 & 1065 & 1809 \\
        \hline
        GMAC & 4055 & 5026 \\
        \hline
        CMAC & 8735 & 8772 \\
        \hline
        KMAC & 11647 & 12186 \\
        \hline
    \end{tabular}
\end{table}

\subsection{KEM performance}
Compared to the $U^{\not\bot}_m$ variant of Fujisaki-Okamoto transformed used in ML-KEM, the encrypt-then-MAC transformation the following trade-off when given the same input sub-routines: \begin{enumerate}
    \item Both encapsulation and decapsulation add a small amount of overhead for needing to hash both the PKE plaintext and the PKE ciphertext when deriving the shared secret, where as the $U^{\not\bot}_m$ transformation only needs to hash the PKE plaintext.
    \item The encapsulation routine adds a small amount of run-time overhead for computing the authenticator
    \item The decapsulation routine enjoys substantial runtime speedup because \emph{re-encryption} is replaced with computing an authenticator
    \item Ciphertext size increases by the size of an authenticator
\end{enumerate}

Since \texttt{K-PKE.\encrypt} carries significantly more computational complexity than \texttt{K-PKE.\decrypt} or any MAC we chose, the performance advantage of the encrypt-then-MAC transformation over the $U^{\not\bot}_m$ transformation is dominated by the runtime saving gained from replacing \emph{re-encryption} with MAC. A comparison between ML-KEM and variations of the $\text{ML-KEM}^+$ can be found in table \ref{tbl:kem-performance}

\begin{table}[h]
    \centering
    \footnotesize
    \caption{CPU cycles of each KEM routine}\label{tbl:kem-performance}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf 128-bit security} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        $\pk$ size & 800 \\
        $\sk$ size & 1632 \\
        \text{ct} size & 768 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 75945 \\
        Average & 76171 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \text{ML-KEM-512} & 91467 & 92065 & 121185 & 121650 \\
        \hline
        \text{Kyber512} & 97811 & 98090 & 119937 & 120299 \\
        \hline
        % ML-KEM+-512
        $\text{ML-KEM}^+$-512 w/ Poly1305 & 93157 & 93626 & 33733 & 33908 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ GMAC & 97369 & 97766 & 37725 & 37831 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ CMAC & 99739 & 99959 & 40117 & 39943 \\
        \hline
        $\text{ML-KEM}^+$-512 w/ KMAC & 101009 & 101313 & 40741 & 40916 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf 192-bit security} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        $\pk$ size & 1184 \\
        $\sk$ size & 2400 \\
        \text{ct} size & 1088 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 129895 \\
        Average & 130650 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \text{ML-KEM-768} & 136405 & 147400 & 186445 & 187529 \\
        \hline
        \text{Kyber768} & 153061 & 153670 & 182129 & 182755 \\
        \hline
        $\text{ML-KEM}^+$-768 w/ Poly1305 & 146405 & 146860 & 43315 & 43463 \\
        \hline
        $\text{ML-KEM}^+$-768 w/ GMAC & 149525 & 150128 & 46513 & 46706 \\
        \hline
        $\text{ML-KEM}^+$-768 w/ CMAC & 153139 & 153735 & 49841 & 50074 \\
        \hline
        $\text{ML-KEM}^+$-768 w/ KMAC & 155219 & 155848 & 52415 & 52611 \\
        \hline
    \end{tabular}\vspace{0.3cm}

    \begin{tabular}[t]{|cc|}
        \hline
        \multicolumn{2}{|c|}{\bf 256-bit security} \\
        \hline
        \multicolumn{2}{|c|}{size parameters (bytes)} \\
        $\pk$ size & 1568 \\
        $\sk$ size & 3168 \\
        \text{ct} size & 1568 \\
        \hline
        \multicolumn{2}{|c|}{$\keygen$ cycles/tick} \\
        Median & 194921 \\
        Average & 195465 \\
        \hline
    \end{tabular}
    \begin{tabular}[t]{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant} 
        & \multicolumn{2}{|c|}{$\encap$ cycles/tick} 
        & \multicolumn{2}{|c|}{$\decap$ cycles/tick} \\
        \cline{2-5}
        & Median & Average & Median & Average \\
        \hline
        \text{ML-KEM-1024} & 199185 & 199903 & 246245 & 247320 \\
        \hline
        \text{Kyber1024} & 222351 & 223260 & 258231 & 259067 \\
        \hline
        $\text{ML-KEM}^+$-1024 w/ Poly1305 & 205763 & 206499 & 51375 & 51562 \\
        \hline
        $\text{ML-KEM}^+$-1024 w/ GMAC & 208805 & 209681 & 54573 & 54780 \\
        \hline
        $\text{ML-KEM}^+$-1024 w/ CMAC & 213667 & 214483 & 59175 & 59408 \\
        \hline
        $\text{ML-KEM}^+$-1024 w/ KMAC & 216761 & 217468 & 62269 & 62516 \\
        \hline
    \end{tabular}
\end{table}

\subsection{Key exchange protocols}\label{sec:key-exchange-protocols}
A common application of key encapsulation mechanism is key exchange protocols, where two parties establish a shared secret using a public channel. \cite{DBLP:conf/eurosp/BosDKLLSSSS18} described three key exchange protocols: unauthenticated key exchange (KE), unilaterally authenticated key exchange (UAKE), and mutually authenticated key exchange (AKE). We instantiated an implementation for each of the three key exchange protocols using different variations of the encrypt-then-MAC KEM and compared round trip time with implementations instantiated using ML-KEM.

For clarity, we denote the party who sends the first message to be the client and the other party to be the server. Round trip time (RTT) is defined to be the time interval between the moment before the client starts generating ephemeral keypairs and the moment after the client derives the final session key. All experiements are run on a pair of AWS c7a.medium instances both located in the \texttt{us-west-2} region. For each experiment, a total of 10,000 rounds of key exchange are performed, with the median and average round trip time (measured in microsecond) recorded.

\subsubsection{Unauthenticated key exchange (KE)}\label{sec:unauthenticated-key-exchange}
In unauthenticated key exchange, a single pair of ephemeral keypair $(\pk_e, \sk_e) \leftsample \keygen()$ is generated by the client. The client transmits the ephemeral public key $\pk_e$ to the server, who runs the encapsulation routine $(c_e, K_e) \leftsample \encap(\pk_e)$ and transmits the ciphertext $c_e$ back to the client. The client finally decapsulates the ciphertext to recover the shared secret $K_e \leftarrow \decap(\sk_e, c_e)$. The key exchange routines are summarized in figure \ref{fig:ke-routines}.

Note that in our implementation, a key derivation function (\texttt{KDF}) is applied to the ephemeral shared secret to derive the final session key. This step is added to maintain consistency with other authenticated key exchange protocols, where the final session key is derived from multiple shared secrets. The key derivation function is instantiated using Shake256, and the final session key is 256 bits in length.

\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{KE}_\texttt{C}()$}\label{alg:ke-client}
            \begin{algorithmic}[1]
                \State $(\pk_e, \sk_e) \leftsample \keygen()$
                \State $\texttt{send}(\pk_e)$
                \State $c_e \leftarrow \texttt{read}()$
                \State $K_e \leftarrow \decap(\sk_e, c_e)$
                \State $K \leftarrow \texttt{KDF}(K)$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[b]{0.49\textwidth}
            \begin{algorithm}[H]
                \caption*{$\texttt{KE}_\texttt{S}()$}\label{alg:ke-server}
                \begin{algorithmic}[1]
                    \State $\pk_e \leftarrow \texttt{read}()$
                    \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                    \State $\texttt{send}(c_e)$
                    \State $K \leftarrow \texttt{KDF}(K_e)$
                    \State \Return $K$
                \end{algorithmic}
            \end{algorithm}
    \end{minipage}

    \caption{Unauthenticated key exchange (KE) routines}\label{fig:ke-routines}
\end{figure}

The RTT comparison is summarized in table \ref{tbl:ke-rtt}

\begin{table}[H]
    \centering
    \small
    \caption{KE RTT comparison}\label{tbl:ke-rtt}

    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 800 & 768 & 92 & 97 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 800 & 784 & 70 & 72 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 800 & 784 & 73 & 76 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 800 & 784 & 75 & 79 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 800 & 784 & 76 & 78 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 1184 & 1088 & 135 & 140 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 1184 & 1104 & 99 & 104 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 1184 & 1104 & 101 & 105 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 1184 & 1104 & 103 & 109 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 1184 & 1104 & 103 & 107 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 1568 & 1568 & 193 & 199 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 1568 & 1584 & 138 & 141 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 1568 & 1584 & 140 & 145 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 1568 & 1584 & 143 & 148 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 1568 & 1584 & 144 & 149 \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Unilaterally authenticated key exchange (UAKE)}\label{sec:uakex}
In unilaterally authenticated key exchange, the authenticating party proves its identity to the other party by demonstrating possession of a secret key that corresponds to a published long-term public key. In this implementation, the client possesses the long-term public key $\pk_S$ of the server, and the server authenticates itself by demonstrating possession of the corresponding long-term secret key $\sk_S$. UAKE routines are summarized in figure \ref{fig:uake-routines}.

In addition to the long-term key, the client will also generate an ephemeral keypair as it does in an unauthenticated key exchange, and the session key is derived by applying the \texttt{KDF} to the concatenation of both the ephemeral shared secret and the shared secret encapsulated under server's long-term key. This helps the key exchange to achieve weak forward secrecy \cite{DBLP:conf/eurocrypt/CanettiK01}.

Using KEM for authentication is especially interesting within the context of post-quantum cryptography: post-quantum KEM schemes usually enjoy better performance characteristics than post-quantum signature schemes with faster runtime, smaller memory footprint, and smaller communication sizes. KEMTLS was proposed in 2020 as an alternative to existing TLS handshake protocols, and many experimental implementations have demonstrated the performance advantage \cite{DBLP:conf/ccs/SchwabeSW20}.

\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{UAKE}_\texttt{C}(\pk_S)$}
            \begin{algorithmic}[1]
                \Require Server's long-term $\pk_S$
                \State $(\pk_e, \sk_e) \leftsample \keygen()$
                \State $(c_S, K_S) \leftsample \encap(\pk_S)$
                \State $\texttt{send}(\pk_e, c_S)$
                \State $c_e \leftarrow \texttt{read}()$
                \State $K_e \leftarrow \decap(\sk_e, c_e)$
                \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}\hfill
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{UAKE}_\texttt{S}(\sk_S)$}
            \begin{algorithmic}[1]
                \Require Server's long-term $\sk_S$
                \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
                \State $K_S \leftarrow \decap(\sk_S, c_S)$
                \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                \State $\texttt{send}(c_e)$
                \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S)$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{Unilaterally authenticated key exchange (UAKE) routines}\label{fig:uake-routines}
\end{figure}

\begin{table}[H]
    \centering
    \small
    \caption{UAKE RTT comparison}\label{tbl:uake-rtt}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 1568 & 768 & 145 & 151 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 784 & 103 & 106 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 784 & 106 & 110 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 784 & 108 & 112 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 784 & 109 & 113 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 2272 & 1088 & 215 & 222 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 1104 & 144 & 150 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 1104 & 149 & 156 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 1104 & 153 & 160 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 1104 & 154 & 159 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 3136 & 1568 & 310 & 318 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 1584 & 202 & 209 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 1584 & 212 & 228 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 1584 & 212 & 218 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 1584 & 213 & 220 \\
        \hline
    \end{tabular}
\end{table}

\subsubsection{Mutually authenticated key exchange (AKE)}\label{sec:akex}
Mutually authenticated key exchange is largely identical to unilaterally authenticated key exchange, except for that client authentication is required. This means that client possesses server's long-term public key and its own long-term secret key, while the server possesses client's long-term public key and its own long-term secret key. The session key is derived by applying \texttt{KDF} onto the concatenation of shared secrets produced under the ephemeral keypair, server's long-term keypair, and client's long-term keypair, in this order.

\begin{figure}[H]
    \centering
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{AKE}_\texttt{C}(\pk_S, \sk_C)$}
            \begin{algorithmic}[1]
                \Require Server's long-term $\pk_S$
                \Require Client's long-term $\sk_C$
                \State $(\pk_e, \sk_e) \leftsample \keygen()$
                \State $(c_S, K_S) \leftsample \encap(\pk_S)$
                \State $\texttt{send}(\pk_e, c_S)$
                \State $(c_e, c_C) \leftarrow \texttt{read}()$
                \State $K_e \leftarrow \decap(\sk_e, c_e)$
                \State $K_C \leftarrow \decap(\sk_e, c_C)$
                \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}
    \begin{minipage}[b]{0.49\textwidth}
        \begin{algorithm}[H]
            \caption*{$\texttt{AKE}_\texttt{S}(\sk_S, \pk_C)$}
            \begin{algorithmic}[1]
                \Require Server's long-term $\sk_S$
                \Require Client's long-term $\pk_C$
                \State $(\pk_e, c_S) \leftarrow \texttt{read}()$
                \State $K_S \leftarrow \decap(\sk_S, c_S)$
                \State $(c_e, K_e) \leftsample \encap(\pk_e)$
                \State $(c_C, K_C) \leftsample \encap(\pk_C)$
                \State $\texttt{send}(c_e, c_C)$
                \State $K \leftarrow \texttt{KDF}(K_e \Vert K_S \Vert K_C)$
                \State \Return $K$
            \end{algorithmic}
        \end{algorithm}
    \end{minipage}

    \caption{Mutually authenticated key exchange (AKE) routines}\label{fig:ake-routines}
\end{figure}

\begin{table}[H]
    \centering
    \small
    \caption{AKE RTT comparison}\label{tbl:ake-rtt}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-512} & 1568 & 1536 & 220 & 213 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ Poly1305 & 1584 & 1568 & 133 & 138 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ GMAC & 1584 & 1568 & 139 & 143 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ CMAC & 1584 & 1568 & 143 & 148 \\
        \hline
        $\texttt{ML-KEM-512}^+$ w/ KMAC & 1584 & 1568 & 145 & 151 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-768} & 2272 & 2176 & 294 & 301 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ Poly1305 & 2288 & 2208 & 190 & 196 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ GMAC & 2288 & 2208 & 197 & 210 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ CMAC & 2288 & 2208 & 202 & 208 \\
        \hline
        $\texttt{ML-KEM-768}^+$ w/ KMAC & 2288 & 2208 & 204 & 210 \\
        \hline
    \end{tabular}\vspace{0.3cm}
    \begin{tabular}{|c|c|c|c|c|}
        \hline
        \multirow{2}{*}{KEM variant}
        & \multirow{2}{*}{Client TX bytes}
        & \multirow{2}{*}{Server TX bytes}
        & \multicolumn{2}{|c|}{RTT time ($\us$)} \\
        \cline{4-5}
        & & & Median & Average \\
        \hline
        \texttt{ML-KEM-1024} & 3136 & 3136 & 512 & 511 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ Poly1305 & 3152 & 3168 & 266 & 273 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ GMAC & 3152 & 3168 & 273 & 282 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ CMAC & 3152 & 3168 & 280 & 287 \\
        \hline
        $\texttt{ML-KEM-1024}^+$ w/ KMAC & 3152 & 3168 & 282 & 288 \\
        \hline
    \end{tabular}
\end{table}

\section{Conclusions and future works}\label{sec:future-works}
The encrypt-then-MAC transformation is a generic KEM construction whose IND-CCA security reduces to the OW-PCA security of the input PKE and the one-time existential unforgeability of the input MAC in the random oracle model. Compared to the Fujisaki-Okamoto transformation, the encrypt-then-MAC replaces the computationally expensive re-encryption with computing a MAC tag. At the cost of minimal increase in encapsulation cost and ciphertext size, the encrypt-then-MAC substantially improves the efficiency of the decapsulation routine. Where the input PKE's encryption is slower than decryption, the encrypt-then-MAC KEM achieves meaningful time savings in practical key exchange protocols.

Unfortunately, ML-KEM is not OW-PCA secure. In fact, as Chris Peikert pointed out in \cite{DBLP:conf/pqcrypto/Peikert14}, most lattice-based cryptosystems are not OW-PCA secure due to the search-decision equivalence of lattice problems. While we instantiated encrypt-then-MAC with ML-KEM subroutines for performance comparison, the resulting KEM is not chosen-ciphertext secure, and should not be used in production systems. The natural questions is thus to find a suitable cryptosystem to instantiate encrypt-then-MAC with.

\textbf{RSA and ElGamal.} Combining encrypt-then-MAC with ElGamal results in the ``Hashed ElGamal'' scheme proposed in \cite{DBLP:journals/iacr/AbdallaBR99}\cite{DBLP:conf/ctrsa/AbdallaBR01}. RSA is also known to be OW-PCA secure because it is a trapdoor permutation and thus a rigid PKE. However, because of RSA's rigidity, there exists even more efficient KEM transformation (such as RSA-KEM \cite{DBLP:journals/iacr/Shoup01}) that only adds a single hash to the base PKE routines. While applying encrypt-then-MAC to RSA will result in an IND-CCA secure KEM, such construction offers no meaningful advantage to RSA-KEM.

\textbf{Code-based cryptography.} Because the general problem of decoding a linear code is proven NP hard \cite{DBLP:journals/tit/BerlekampMT78}, code-based cryptosystems may not suffer from the inherent search-decision equivalence of lattice problems and thus be viable candidates with OW-PCA security. Unfortunately, among the code-based submissions to NIST PQC, HQC \cite{melchor2018hamming} and BIKE \cite{aragon2022bike} are known to be vulnerable to key-recovery plaintext-checking attacks (KR-PCA) \cite{DBLP:journals/tches/TanakaUXITH23}. On the other hand, classic McEliece \cite{classicmceliecespec} seems to be PCA secure and thus a viable candidate, though in classic McEliece, the decoding routine is more expensive than the encryption routine, so applying encrypt-then-MAC may not yield meaningful performance gains.

\textbf{Isogeny-based cryptography.} The intractability assumptions of isogeny-based cryptography resemble the classical Diffie-Hellman assumptions, and it seems possible to formulate a ``Gap Diffie-Hellman assumption'' in supersingular isogeny \cite{DBLP:conf/icisc/FujiokaTTY18}. While SIKE and SIDH were found to be insecure \cite{DBLP:conf/eurocrypt/CastryckD23}, other isogeny-based cryptosystem such as CSIDH \cite{DBLP:conf/asiacrypt/Castryck0MPR18} remains unaffected by the aforementioned attack and might be suitable candidates.

%%%% 8. BILBIOGRAPHY %%%%
\bibliographystyle{alpha}
\bibliography{abbrev3,crypto,biblio}
% - Download abbrev3.bib and crypto.bib from https://cryptobib.di.ens.fr/
% - Use bilbio.bib for additional references not in the cryptobib database.
%   If possible, take them from DBLP.

% \section{Appendix}\label{sec:appendix}
% % TODO: brief introduction to the various authenticators should be in the appendix

% \subsection{Performance comparison between $\texttt{ML-KEM}^+$ and Kyber}\label{sec:ml-kem-and-kyber}
% ML-KEM directly evolved from CRYSTALS-Kyber's third round submission to NIST's post quantum cryptography competition. While their IND-CPA subroutines (see figure \ref{fig:k-pke-routines}) are identical, ML-KEM deviated from Kyber by choosing a different variant of the Fujisaki-Okamoto transformation.

% \begin{figure}[H]
%     \centering

%     \begin{minipage}[t]{0.33\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.KeyGen()}}
%             \begin{algorithmic}[1]
%                 \State $A \leftsample R_q^{k \times k}$
%                 \State $\mathbf{s} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{e} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{t} \leftarrow A\mathbf{s} + \mathbf{e}$
%                 \State $\pk \leftarrow (A, \mathbf{t})$
%                 \State $\sk \leftarrow \mathbf{s}$
%                 \State \Return $(\pk, \sk)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{0.33\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.Enc(pk, m)}}
%             \begin{algorithmic}[1]
%                 \Ensure $\pk = (A, \mathbf{t})$
%                 \Ensure $m \in R_2$
%                 \State $\mathbf{r} \leftsample \mathcal{X}_{\eta_1}^k$
%                 \State $\mathbf{e}_1 \leftsample \mathcal{X}_{\eta_2}^k$
%                 \State $e_2 \leftsample \mathcal{X}_{\eta_2}$
%                 \State $\mathbf{c}_1 \leftarrow A\mathbf{r} + \mathbf{e}_1$
%                 \State $c_2 \leftarrow \mathbf{t}^\intercal \mathbf{r} + e_2 + m \cdot \lfloor \frac{q}{2}\rceil$
%                 \State \Return $(\mathbf{c}_1, c_2)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{0.3\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{K-PKE.Dec(sk, c)}}
%             \begin{algorithmic}[1]
%                 \Ensure $c = (\mathbf{c}_1, c_2)$
%                 \Ensure $\sk = \mathbf{s}$
%                 \State $\hat{m} \leftarrow c_2 - \mathbf{c}_1^\intercal \cdot \mathbf{s}$
%                 \State $\hat{m} \leftarrow \texttt{Round}(\hat{m})$
%                 \State \Return $\hat{m}$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}
%     \caption{\texttt{K-PKE} routines are identical between Kyber and ML-KEM}\label{fig:k-pke-routines}
% \end{figure}

% CRYSTALS-Kyber uses the $U^{\not\bot}$ variant, where the shared secret is derived from both the plaintext and the ciphertext. On the other hand, because by using \emph{re-encryption} and \emph{de-randomization}, the PKE is already made \emph{rigid}, the CRYSTALS-Kyber team decided to use the $U^{\not\bot}_m$ variant, where the shared secret is derived from the plaintext alone.

% \begin{figure}[H]
%     \centering
%     \begin{minipage}[t]{0.45\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.KeyGen()}}
%             \begin{algorithmic}[1]
%                 \State $z \leftsample \{0,1\}^{256}$
%                 \State $(\pk, \sk^\prime) \leftsample \texttt{PKE.KeyGen}()$
%                 \State $\sk \leftarrow (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
%                 \State \Return $(\pk, \sk)$
%             \end{algorithmic}
%         \end{algorithm}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.Encap(pk)}}
%             \begin{algorithmic}[1]
%                 \State $m \leftsample \{0,1\}^{256}$
%                 \State $(\overline{K}, r) \leftarrow G(m \Vert H(\pk))$
%                 \State $c \leftarrow \texttt{PKE.Enc}(\pk, m, r)$
%                 \State $K \leftarrow \texttt{KDF}(\overline{K} \Vert H(c))$
%                     \Comment{$U^{\not\bot}$}
%                 \State $K \leftarrow \overline{K}$
%                     \Comment{$U^{\not\bot}_m$}
%                 \State \Return $(c, K)$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}
%     \begin{minipage}[t]{0.5\textwidth}
%         \begin{algorithm}[H]
%             \caption*{\texttt{KEM.Decap(sk, c)}}
%             \begin{algorithmic}[1]
%                 \Ensure $\sk = (\sk^\prime \Vert \pk \Vert H(\pk) \Vert z)$
%                 \State $\hat{m} \leftarrow \texttt{PKE.Dec}(\sk^\prime, c)$
%                 \State $(\overline{K}, \hat{r}) \leftarrow G(\hat{m} \Vert H(\pk))$
%                 \If{$\texttt{PKE.Enc}(\pk, \hat{m}, \hat{r}) = c$}
%                     \State $K \leftarrow \texttt{KDF}(\overline{K}, H(c))$
%                         \Comment{$U^{\not\bot}$}
%                     \State $K \leftarrow \overline{K}$
%                         \Comment{$U^{\not\bot}_m$}
%                 \Else 
%                     \State $K \leftarrow \texttt{KDF}(z \Vert H(c))$
%                 \EndIf
%                 \State \Return $K$
%             \end{algorithmic}
%         \end{algorithm}
%     \end{minipage}

%     \caption{
%         Kyber uses $U^{\not\bot}$ variant. ML-KEM uses $U^{\not\bot}_m$ variant.
%     }\label{fig:kyber-ml-kem-compare}
% \end{figure}

% The reason for ML-KEM to use a different variant of the Fujisaki-Okamoto transformation is two-fold. The first reason is performance: using the $U^{\not\bot}_m$ transformation saves the need to hash the ciphertext, and since Kyber/ML-KEM's performance is mainly bottlenecked by the symmetric components, omitting the hash leads to significant runtime savings (up to 17\% in AVX-2 optimized implementations). The second reason is the simplified security proof and tighter security bounds of the $U^{\not\bot}_m$ variant compared to the $U^{\not\bot}$ variant. We will omit the details of the security proof and refer readers to \cite{hofheinz2017modular}.

% In section \ref{sec:implementation}, we mainly compared $\texttt{ML-KEM}^+$ with \texttt{ML-KEM}, but the we would like to point out that, because Kyber uses the $U^{\not\bot}$ variant and needs to hash the ciphertext for deriving the shared secret, the performance advantage of $\texttt{ML-KEM}^+$ over Kyber will be even greater.

\end{document}